### 2023-06-15 assignment-4/article.tex
Get bibliography working

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Arsclassica Article
% LaTeX Template
% Version 1.1 (1/8/17)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Lorenzo Pantieri (http://www.lorenzopantieri.net) with extensive modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
10pt, % Main document font size
a4paper, % Paper type, use 'letterpaper' for US Letter paper
oneside, % One page layout (no page indentation)
%twoside, % Two page layout (page indentation for binding and different headers)
headinclude,footinclude, % Extra spacing for the header and footer
BCOR5mm, % Binding correction
]{scrartcl}

\input{structure.tex} % Include the structure.tex file which specified the document structure and layout

\hyphenation{Fortran hy-phen-ation} % Specify custom hyphenation points in words with dashes where you would like hyphenation to occur, or alternatively, don't put any dashes in a word to stop hyphenation altogether

%----------------------------------------------------------------------------------------
%	TITLE AND AUTHOR(S)
%----------------------------------------------------------------------------------------

\title{\normalfont\spacedallcaps{Language reclamation: Literature review}} % The article title

%\subtitle{Subtitle} % Uncomment to display a subtitle

\author{\spacedlowsmallcaps{Caleb Moses*}} % The article author(s) - author affiliations need to be specified in the AUTHOR AFFILIATIONS block

\date{} % An optional date to appear under the author(s)

%----------------------------------------------------------------------------------------

\begin{document}

\title{MATH/COMP 562 ASSIGNMENT 4}
\date{Due: April 9th (Sunday)}
\author{Adam M. Oberman}
\maketitle

\section*{Generative Models}
Refer to \url{https://udlbook.github.io/udlbook/}

\subsection*{Exercise 4.1 (GANs)}

Given a dataset of genuine images, $S_m = \{x_1, \ldots, x_m\}$. The goal is to train a generator, $x_i = g(z_i, \theta)$, to generate images $x_i \in S_m$ from noise $z_i$. This is achieved by also training a discriminator, $f(x, \phi)$, which classifies an image as genuine or generated, using the minimax training loss
\begin{equation*}
\max_\theta \min_\phi \mathcal{L}^S_\theta(f(x, \phi), y)
\end{equation*}
applied to the combined dataset, $S_\theta = S_N \cup S_G^\theta$,
where
\begin{itemize}
    \item $S_N = \{(x_1, 1), \ldots, (x_m, 1)\}$, the (fixed) genuine images, with label $y = 1$,
    \item $S_G^\theta = \{(x^\theta_1, 0), \ldots, (x^\theta_m, 0)\}$, a changing set of generated images $x^\theta_i = g(z_i, \theta)$, with label $y = 0$.
\end{itemize}

\subsubsection*{Explicit GAN loss}

From \citet{prince2023understanding} the GAN loss function is as follows

%% \begin{equation*}
%% \hat{\theta} = \textrm{argmax}_\theta \left[ \min_\phi \left[ \sum \right] \right]
%% \end{equation*}

\subsubsection*{Exercise 4.1 (a)}
Write down and explain the loss for the discriminator, if the generator is fixed.

\subsubsection*{Solution 4.1 (a)}
For a fixed generator, we can imagine

Assuming the generator is fixed, the discriminator's goal is to correctly classify the genuine images and the generated images. The loss for the discriminator is a binary cross-entropy loss function, which measures the error of a classification model at two-class classification tasks. This loss increases as the predicted probability diverges from the actual label. The binary cross-entropy loss can be expressed as follows:

\begin{equation}
L(y,\hat{y}) = - \frac{1}{N} \sum_{i=1}^{N} [y_i \log(\hat{y_i}) + (1-y_i) \log(1-\hat{y_i})]
\end{equation}

where $y$ is the true label, $\hat{y}$ is the predicted label, and $N$ is the number of samples.

\subsubsection*{Exercise 4.1 (b)}
Write down and explain the loss (or gain, since it is a maximization) for the generator, if the discriminator is fixed.

\subsubsection*{Solution 4.1 (b)}
The generator's loss function is also the binary cross-entropy loss function. However, the goal is the opposite, i.e., it tries to maximize this loss. The generator wants the discriminator to incorrectly classify its generated images as genuine. Thus, when the discriminator is fixed, the generator tries to generate images that look more like the genuine images to trick the discriminator.

\subsubsection*{Exercise 4.1 (c)}
Characterize the saddle points of the loss. If the generator outputs the full set of genuine images, is this optimal? What if the generator outputs just a subset of the genuine images?

A saddle point, $(\theta, \phi)$, is a point where (i) if we fix $\phi$ and change $\theta$, we cannot further increase the loss, and (ii) if we fix $\theta$ and change $\phi$, we cannot further decrease the loss.

\subsubsection*{Solution 4.1 (c)}
Saddle points are the points where the gradient of the loss function equals zero, i.e., the function is neither entirely convex nor entirely concave. If the generator outputs the full set of genuine images, this is not optimal, as the generator is not adding any new information. Similarly, if the generator only outputs a subset of the genuine images, it may not be optimizing the diversity of the generated images, which is also not optimal.

\subsection*{Exercise 4.2 (Diffusion generative models)}
Given the dataset $S_m = \{x_1 , \ldots , x_m \}$, the goal is to generate images from the dataset. The inputs are a finite set of numbers $\alpha_t , t \in T$ and a finite set of noise vectors $\zeta_j , j \in J$.

The outputs are a set of models, $g_t : X \rightarrow X$, $g_t (x) = g(x, t), t \in T$.

For each $t \in T$ , let $S_t = \{x_{t,i,j}\}, i = 1, \ldots , m, j \in J$ be the dataset of t-noisy images, $x_{t,i,j} = \sqrt{\alpha_t} x_i + \sqrt{1 - \alpha_t} \zeta_j$

The training loss for $g_t$ is given by
\begin{equation*}
\mathcal{L} S_t (g_t ) = \frac{1}{m} \sum_{i=1}^{m} \sum_{j \in J} ||g_t (x_{t,i,j} ) - \zeta_j ||^2
\end{equation*}

Define the closest point map
\begin{equation*}
g_{CP}(x) = g_{CP} (x, S_m ) = \arg \min_{x_i \in S_m} ||x - x_i ||^2
\end{equation*}

\subsubsection*{Exercise 4.2 (a)}
Assume that, for each $x_{t,i,j} \in S_t$, $g_{CP}(x_{t,i,j}) = x_i$. Find the loss minimizer $g^*_t$ and show that the loss is zero.

\subsubsection*{Solution to Exercise 4.2 (a)}
This requires a specific calculation or a specific dataset to answer more accurately. The answer might be based on the characteristics of the dataset and the model's architecture.

\subsubsection*{Exercise 4.2 (b)}
In the case above, are there any other loss minimizers?

\subsubsection*{Solution to Exercise 4.2 (b)}
This question also requires more specific details about the dataset or the model's architecture to provide a complete and accurate answer.

\subsubsection*{Exercise 4.2 (c)}
How do the loss minimizers for this generative model compare to the loss minimizers for GANs? Which is better in terms of coverage of the dataset?

\subsubsection*{Solution to Exercise 4.2 (c)}
The comparison between the loss minimizers for this generative model and the loss minimizers for GANs is largely dependent on the specific characteristics of the models and the dataset.

\subsection*{Exercise 4.3 (Normalizing Flows: transformation of densities)}
Problems 16.1 and 16.2 from ”Understanding Deep Learning”

\subsection*{Exercise 4.4 (Normalizing Flows: Fixed point theorem)}
Problem 16.11 from Chapter 16 of ”Understanding Deep Learning”

\subsection*{Exercise 4.5 (Diffusion generative models)}
Problems 18.1 and 18.2 from Chapter 18 of ”Understanding Deep Learning”

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\bibliography{sample.bib} % The file containing the bibliography

\end{document}

### 2023-06-15 assignment-4/structure.tex
Get bibliography working

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Arsclassica Article
% Structure Specification File
%
% This file has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Lorenzo Pantieri (http://www.lorenzopantieri.net) with extensive modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	REQUIRED PACKAGES
%----------------------------------------------------------------------------------------

\usepackage[
nochapters, % Turn off chapters since this is an article        
beramono, % Use the Bera Mono font for monospaced text (\texttt)
eulermath,% Use the Euler font for mathematics
pdfspacing, % Makes use of pdftex’ letter spacing capabilities via the microtype package
dottedtoc % Dotted lines leading to the page numbers in the table of contents
]{classicthesis} % The layout is based on the Classic Thesis style

\usepackage{arsclassica} % Modifies the Classic Thesis package

\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs

\usepackage[utf8]{inputenc} % Required for including letters with accents

\usepackage{graphicx} % Required for including images
\graphicspath{{Figures/}} % Set the default folder for images

\usepackage{enumitem} % Required for manipulating the whitespace between and within lists

\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template

\usepackage{subfig} % Required for creating figures with multiple parts (subfigures)

\usepackage{amsmath,amssymb,amsthm} % For including math equations, theorems, symbols, etc

\usepackage[round]{natbib}   % omit 'round' option if you prefer square brackets

\usepackage{varioref} % More descriptive referencing

\bibliographystyle{plainnat}

%----------------------------------------------------------------------------------------
%	THEOREM STYLES
%---------------------------------------------------------------------------------------

\theoremstyle{definition} % Define theorem styles here based on the definition style (used for definitions and examples)
\newtheorem{definition}{Definition}

\theoremstyle{plain} % Define theorem styles here based on the plain style (used for theorems, lemmas, propositions)
\newtheorem{theorem}{Theorem}

\theoremstyle{remark} % Define theorem styles here based on the remark style (used for remarks and notes)

%----------------------------------------------------------------------------------------
%	HYPERLINKS
%---------------------------------------------------------------------------------------

\hypersetup{
%draft, % Uncomment to remove all links (useful for printing in black and white)
colorlinks=true, breaklinks=true, bookmarks=true,bookmarksnumbered,
urlcolor=webbrown, linkcolor=RoyalBlue, citecolor=webgreen, % Link colors
pdftitle={}, % PDF title
pdfauthor={\textcopyright}, % PDF Author
pdfsubject={}, % PDF Subject
pdfkeywords={}, % PDF Keywords
pdfcreator={pdfLaTeX}, % PDF Creator
pdfproducer={LaTeX with hyperref and ClassicThesis} % PDF producer
}

### 2023-06-18 assignment-4/article.tex
14.a complete

@@
 
-\subsubsection*{Explicit GAN loss}
+%% \subsubsection*{Explicit GAN loss}
 
-From \citet{prince2023understanding} the GAN loss function is as follows
+%% From \citet{prince2023understanding} the GAN loss function is as follows:
 
+%% %% This is the characterisation of the saddle points
 %% \begin{equation*}
-%% \hat{\theta} = \textrm{argmax}_\theta \left[ \min_\phi \left[ \sum \right] \right]
+%% \hat{\theta} = \textrm{argmax}_\theta \left[ \min_\phi \left[ \sum_j - \log(1 - D(G(z_j, \theta), \phi))) - \sum_i \log(D(x_i, \phi))) \right] \right]
 %% \end{equation*}
+
+%% where $\theta$ represents the weights of the generator function $G$, and $\phi$ is the weights of the discriminator $D$, and $\sigma$ is the sigmoid logistic function.
 
@@
 \subsubsection*{Solution 4.1 (a)}
-For a fixed generator, we can imagine
 
-Assuming the generator is fixed, the discriminator's goal is to correctly classify the genuine images and the generated images.
-The loss for the discriminator is a binary cross-entropy loss function, which measures the error of a classification model at two-class classification tasks.
-This loss increases as the predicted probability diverges from the actual label.
-The binary cross-entropy loss can be expressed as follows:
+From \citet{prince2023understanding} 15.5 the discriminator loss function is given by:
 
-\begin{equation}
-L(y,\hat{y}) = - \frac{1}{N} \sum_{i=1}^{N} [y_i \log(\hat{y_i}) + (1-y_i) \log(1-\hat{y_i})]
-\end{equation}
+\begin{equation*}
+  L(\phi) = \sum_j - \log\left(1 - \textrm{sig}(f(g(z_j, \theta), \phi)\right) - \sum_i log\left(\textrm{sig}(f(x_i, \phi))\right)
+\end{equation*}
 
-where $y$ is the true label, $\hat{y}$ is the predicted label, and $N$ is the number of samples.
+where $\textrm{sig}$ represents the sigmoid function.
+
+For a fixed generator, we can imagine each $g(z_i, \theta) = \hat{x}_i$ where $\hat{x}_i$ is an image generated according to $g$ which approximates $x_i$.
+We can simplify further if we let $D(x, \phi) = \textrm{sig}(f(x, \phi))$, basically considering the sigmoid classification layer as part of the discriminator function.
+
+This gives us the loss function below:
+
+\begin{equation*}
+  L(\phi) = \sum_j - \log(1 - D(\hat{x}_j, \phi)) - \sum_i log(D(x_i, \phi))
+\end{equation*}
+
+The result is a binary cross-entropy loss where the goal is to distinguish $x_i$ from $\hat{x}_i$.
 
@@
 \subsubsection*{Solution 4.1 (b)}
-The generator's loss function is also the binary cross-entropy loss function.
-However, the goal is the opposite, i.e., it tries to maximize this loss.
-The generator wants the discriminator to incorrectly classify its generated images as genuine.
-Thus, when the discriminator is fixed, the generator tries to generate images that look more like the genuine images to trick the discriminator.
+
+For a fixed discriminator, we can ignore the $\min$ part of the GAN loss - for example by assuming $\phi$ is already optimal.
+We can think of $D(x, \phi)$ as $D_\phi(x)$ (we are bringing the $\phi$ into the function, since now it can't vary).
+In this scenario the GAN problem requires us to find the optimal $\hat{\theta}$ satisfying below:
+
+\begin{equation*}
+  \sum_j \log(1 - D_{\phi}(G(z_j, \theta)))
+\end{equation*}
+
+So we are trying to maximise the binary cross-entropy loss of the classification task. The generator wants the discriminator to incorrectly classify the generated images as genuine.
 

### 2023-06-18 assignment-4/article.tex
Whitespace cleanup
4.1a answer refined
4.1b answer complete

@@
 \begin{equation*}
-  L(\phi) = \sum_j - \log\left(1 - \textrm{sig}(f(g(z_j, \theta), \phi)\right) - \sum_i log\left(\textrm{sig}(f(x_i, \phi))\right)
+  L(\phi) = - \sum_j \log\left(1 - \textrm{sig}(f(g(z_j, \theta), \phi)\right) - \sum_i log\left(\textrm{sig}(f(x_i, \phi))\right)
 \end{equation*}
 
-where $\textrm{sig}$ represents the sigmoid function.
+where ``$\textrm{sig}$'' represents the sigmoid function.
 
@@
 \begin{equation*}
-  L(\phi) = \sum_j - \log(1 - D(\hat{x}_j, \phi)) - \sum_i log(D(x_i, \phi))
+  L(\phi) = - \sum_j \log(1 - D(\hat{x}_j, \phi)) - \sum_i log(D(x_i, \phi))
 \end{equation*}
@@
 The result is a binary cross-entropy loss where the goal is to distinguish $x_i$ from $\hat{x}_i$.
+Real samples $x_i$ decrease the loss by $log(D(x_i, \phi))$ while generated samples increase the loss by $\log(1-D(\hat{x}_j, \phi))$.
 
@@
 
-For a fixed discriminator, we can ignore the $\min$ part of the GAN loss - for example by assuming $\phi$ is already optimal.
-We can think of $D(x, \phi)$ as $D_\phi(x)$ (we are bringing the $\phi$ into the function, since now it can't vary).
+From \citet{prince2023understanding} 15.5 the generator gain function is given by:
+
+\begin{equation*}
+  L(\theta) = - \sum_j \log\left(1 - \textrm{sig}(f(g(z_j, \theta), \phi)\right)
+\end{equation*}
+
+where, as before ``$\textrm{sig}$'' represents the sigmoid function.
+
+For a fixed discriminator, we can let $D_\phi(x) = \textrm{sig}(f(x, \phi))$ (we are bringing the $\phi$ into the function, since now it can't vary).
 In this scenario the GAN problem requires us to find the optimal $\hat{\theta}$ satisfying below:
@@
 \begin{equation*}
-  \sum_j \log(1 - D_{\phi}(G(z_j, \theta)))
+  L(\theta) = - \sum_j \log(1 - D_{\phi}(g(z_j, \theta)))
 \end{equation*}
 
-So we are trying to maximise the binary cross-entropy loss of the classification task. The generator wants the discriminator to incorrectly classify the generated images as genuine.
+The purpose of the generator gain function is to reward the generator for samples that are assigned high probability by the discriminator.
 

### 2023-06-18 assignment-4/article.tex
4.1c solved

@@
 ]{scrartcl}
+
+\usepackage{amsmath}
+\DeclareMathOperator*{\argmax}{argmax} % thin space, limits underneath in displays
+\DeclareMathOperator*{\sig}{sig}
 
@@
 \subsubsection*{Solution 4.1 (c)}
-Saddle points are the points where the gradient of the loss function equals zero, i.e., the function is neither entirely convex nor entirely concave.
-If the generator outputs the full set of genuine images, this is not optimal, as the generator is not adding any new information.
-Similarly, if the generator only outputs a subset of the genuine images, it may not be optimizing the diversity of the generated images, which is also not optimal.
+
+The saddle points of the GAN loss function are solutions to the following minimax problem, as described in \citet{prince2023understanding} equation 15.4:
+
+\begin{equation*}
+\hat{\theta} = \argmax_\theta \left[ \min_\phi \left[ \sum_j - \log(1 - \sig(f(g(z_j, \theta), \phi))) - \sum_i \log(\sig(f(x_i, \phi))) \right] \right]
+\end{equation*}
+
+This represents the Nash equilibrium in the minimax game between the generator and the discriminator.
+
+If the generator produces the full set of genuine images, the discriminator could learn to memorize the genuine images and achieve perfect discrimination.
+In this scenario the generator would not be optimal.
+
+If however the generator only produces a subset of the genuine images, then it is sub-optimal.
+This phenomenon is referred to as as "mode collapse", where the generator learns to generate only a limited variety of samples.
+
+\newpage
 

### 2023-06-18 assignment-4/article.tex
4.2a and 4.2b solved

@@
 \DeclareMathOperator*{\argmax}{argmax} % thin space, limits underneath in displays
+\DeclareMathOperator*{\argmin}{argmin} % thin space, limits underneath in displays
 \DeclareMathOperator*{\sig}{sig}
@@
 Given the dataset $S_m = \{x_1 , \ldots , x_m \}$, the goal is to generate images from the dataset.
-The inputs are a finite set of numbers $\alpha_t , t \in T$ and a finite set of noise vectors $\zeta_j , j \in J$.
+The inputs are a finite set of numbers $\alpha_t , t \in T$ and a finite set of noise vectors $\epsilon_j$ for $j \in J$.
 
-The outputs are a set of models, $g_t : X \rightarrow X$, $g_t (x) = g(x, t), t \in T$.
+The outputs are a set of models, $g_t : \mathcal{X} \rightarrow \mathcal{X}$, $g_t (x) = g(x, t)$ for $t \in T$.
 
-For each $t \in T$ , let $S_t = \{x_{t,i,j}\}, i = 1, \ldots , m, j \in J$ be the dataset of t-noisy images, $x_{t,i,j} = \sqrt{\alpha_t} x_i + \sqrt{1 - \alpha_t} \zeta_j$
+For each $t \in T$ , let $S_t = \{x_{t,i,j}\}, i = 1, \ldots , m$ for $j \in J$ be the dataset of t-noisy images, $x_{t,i,j} = \sqrt{\alpha_t} x_i + \sqrt{1 - \alpha_t} \epsilon_j$
 
@@
 \begin{equation*}
-\mathcal{L} S_t (g_t ) = \frac{1}{m} \sum_{i=1}^{m} \sum_{j \in J} ||g_t (x_{t,i,j} ) - \zeta_j ||^2
+\hat{L}_{S_t}(g_t) = \frac{1}{m} \sum_{i=1}^{m} \sum_{j \in J} ||g_t (x_{t,i,j} ) - \epsilon_j ||^2
 \end{equation*}
@@
 \begin{equation*}
-g_{CP}(x) = g_{CP} (x, S_m ) = \arg \min_{x_i \in S_m} ||x - x_i ||^2
+  g^{CP}(x) = g^{CP}(x, S^m) = \argmin_{x_i \in S^m} ||x - x_i||^2
 \end{equation*}
@@
 \subsubsection*{Exercise 4.2 (a)}
-Assume that, for each $x_{t,i,j} \in S_t$, $g_{CP}(x_{t,i,j}) = x_i$. Find the loss minimizer $g^*_t$ and show that the loss is zero.
+Assume that, for each $x_{t,i,j} \in S_t$, $g^{CP}(x_{t,i,j}) = x_i$. Find the loss minimizer $g^*_t$ and show that the loss is zero.
 
 \subsubsection*{Solution to Exercise 4.2 (a)}
-This requires a specific calculation or a specific dataset to answer more accurately. The answer might be based on the characteristics of the dataset and the model's architecture.
+In this argument we will show that under the given assumption, the noise vector $\epsilon_j$ can be retrieved from $x_{t,i,j}$, and this fact can be used to construct a loss minimizing function.
+
+The training loss is the mean squared error between the output of $g_t(x_{t,i,j})$ and the noise vectors $\epsilon_j$:
+
+\begin{equation*}
+  \hat{L}_{S_t}(g_t) = \frac{1}{m} \sum_{i=1}^{m} \sum_{j \in J} ||g_t (x_{t,i,j} ) - \epsilon_j ||^2
+\end{equation*}
+
+Minimizing this loss amounts to minimizing $||g_t (x_{t,i,j})-\epsilon_j||^2$ for every $i = 1, ..., m$, and every $j \in J$.
+
+Since $x_{t,i,j} = \sqrt{\alpha_t}x_i + \sqrt{1-\alpha_t}\epsilon_j$, rearranging for $\epsilon_j$ we can construct a new function $g^*$ which will depend only on $x_{t,i,j}$ and $t$:
+
+\begin{align*}
+  \epsilon_j &= \frac{x_{t,i,j} - \sqrt{\alpha_t}x_i}{\sqrt{1-\alpha_t}} \\
+   &= \frac{x_{t,i,j} - \sqrt{\alpha_t} \cdot \argmin_{x_k \in S^m} ||x_{t,i,j}-x_k||^2}{\sqrt{1-\alpha_t}} \\
+  &= g^*_t(x_{t,i,j})
+\end{align*}
+
+Where we take advantage of the fact that $g^{CP}(x_{t,i,j}) = x_i$ to remove the dependency on $x_i$.
+By construction $g$ depends only on $x_{t,i,j}$ and $t$.
+This means $g^*_t(x)$ is a viable candidate minimizer for the loss $\hat{L}_{S_t}$.
+
+Lastly we show that $g^*_t$ minimizes $\hat{L}_{S_t}$:
+
+\begin{align*}
+  \hat{L}_{S_t}(g^*_t) &= \frac{1}{m} \sum_{i=1}^{m} \sum_{j \in J} ||g^*_t (x_{t,i,j} ) - \epsilon_j ||^2 \\
+  &= \frac{1}{m} \sum_{i=1}^{m} \sum_{j \in J} ||\epsilon_j - \epsilon_j ||^2 \\
+  &= \frac{1}{m} \sum_{i=1}^{m} \sum_{j \in J} 0 = 0
+\end{align*}
+
+Since the loss is non-negative, we can conclude that $g^*_t$ minimizes the loss.
 
@@
 \subsubsection*{Solution to Exercise 4.2 (b)}
-This question also requires more specific details about the dataset or the model's architecture to provide a complete and accurate answer.
+
+Suppose there are two different minimizers $g^t$ and $h^t$ of the loss. They both achieve the minimum loss, which is zero. So for all $i = 1, ..., m$ and for all $j \in J$,
+
+\begin{equation*}
+  ||g^t(x_{t,i,j}) - \epsilon_j||^2 = ||h^t(x_{t,i,j}) - \epsilon_j||^2 = 0
+\end{equation*}
+
+This implies that $g^t(x{t,i,j}) = h^t(x{t,i,j}) = \epsilon_j$ for all $i$ and $j$.
+
+Therefore, if $g^t$ and $h^t$ are both loss minimizers for $\hat{L}_{S^t}$, they must be equal. Therefore $g^*_t$ is unique.
 

### 2023-06-18 assignment-4/article.tex
4.2b simplified slightly
Solve 4.2 c

@@
 
-This implies that $g^t(x{t,i,j}) = h^t(x{t,i,j}) = \epsilon_j$ for all $i$ and $j$.
-
-Therefore, if $g^t$ and $h^t$ are both loss minimizers for $\hat{L}_{S^t}$, they must be equal. Therefore $g^*_t$ is unique.
+This implies that $g^t(x_{t,i,j}) = h^t(x_{t,i,j}) = \epsilon_j$ for all $i$ and $j$.
+Therefore, if $g^t$ and $h^t$ are both loss minimizers for $\hat{L}_{S^t}$, they must be equal.
+Therefore $g^*_t$ is unique.
 
@@
 \subsubsection*{Solution to Exercise 4.2 (c)}
-The comparison between the loss minimizers for this generative model and the loss minimizers for GANs is largely dependent on the specific characteristics of the models and the dataset.
+
+In GANs, the generator aims to fool the discriminator by producing outputs that are indistinguishable from the real data, thus the loss minimizer corresponds to a generator that perfectly replicates the real data distribution.
+However, GANs can suffer from mode collapse, where they only generate a small subset of the real data, failing to cover the entire data distribution.
+
+In contrast, diffusion generative add noise to the real data and learns to reverse this process, mapping noised images back to the originals.
+In this case, the loss minimizer is a deterministic function that directly maps the noisy data back to the closest point in the original dataset.
+This approach virtually guarantees coverage over the original data, as long as the original images are not too close together.
+
+In terms of dataset coverage, diffusion generative models generally perform better. They more directly address mode collapse, which is a notable issue with GANs.
+
+\newpage
 

### 2023-06-18 assignment-4/article.tex
Solve 4.3

@@
 \subsection*{Exercise 4.3 (Normalizing Flows: transformation of densities)}
-Problems 16.1 and 16.2 from ”Understanding Deep Learning”
+
+All problems in this section are drawn from ``Understanding Deep Learning'' \citet{prince2023understanding}.
+
+\subsubsection*{Problem 16.1}
+
+Consider transforming a uniform base density defined on $z\in[0,1]$ using the function $x = f[z] = z^2$. Find an expression for the transformed distribution $Pr(x)$.
+
+\subsubsection*{Problem 16.1 solution}
+
+Let $P_Z$ be the probability density of $z$, then $P_Z = 1$, since $z\sim U[0,1]$.
+
+Now let $x = z^2$, then $z = \sqrt{x}$, and $\frac{dz}{dx} = \frac{1}{2\sqrt{x}}$. Let $P_X$ be the probability density of $x$. Lastly we use the change of variables formula for probability densities:
+
+\begin{align*}
+  p_X(x) &= p_Z(z) \left|\frac{dz}{dx}\right| \\
+  &= 1 \cdot \left| \frac{1}{2 \sqrt{x}} \right| = \frac{1}{2\sqrt{x}}
+\end{align*}
 

### 2023-06-18 assignment-4/article.tex
Solve 4.3
Add the questions for the rest to be solved

@@
 
+\subsubsection*{Problem 16.2}
+
+Consider transforming a standard normal distribution:
+
+\begin{equation*}
+  Pr(z) = \frac{1}{\sqrt{2\pi}} \exp \left[\frac{-z^2}{2}\right]
+\end{equation*}
+
+with the function:
+
+\begin{equation*}
+  x = f[z] = \frac{1}{1 + \exp[-z]}
+\end{equation*}
+
+Find an expression for the transformed distribution $Pr(x)$.
+
+\subsubsection*{Problem 16.2 solution}
+
+As before, let $P_Z = Pr(z)$. Then $x = f[z]$ implies $z = \log(x(1-x))$, from which we derive: $\frac{dz}{dx} = \frac{1-2x}{x(1-x)}$
+
+Then using the change of variables formula from before:
+
+\begin{align*}
+  p_X(x) &= p_Z(z) \left|\frac{dz}{dx}\right| \\
+  &= \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{1}{2}(\log(x(1-x)))^2 \right] \cdot \frac{1-2x}{x(1-x)} \\
+  &= \frac{1}{\sqrt{2\pi}} \exp \left[\log(x(1-x)))^{-\frac{1}{2} \cdot 2} \right] \cdot \frac{1-2x}{x(1-x)} \\
+  &= \frac{1}{\sqrt{2\pi}} \exp \left[\log\left(\frac{1}{x(1-x)}\right) \right] \cdot \frac{1-2x}{x(1-x)} \\
+  &= \frac{1}{\sqrt{2\pi}} \frac{1}{x(1-x)} \cdot \frac{1-2x}{x(1-x)} \\
+  &= \frac{1}{\sqrt{2\pi}} \frac{1-2x}{x^2(1-x)^2}
+\end{align*}
+
+%% \subsubsection*{Problem 16.2 solution - attempt 2}
+
+%% Given the standard normal distribution:
+%% \begin{equation*}
+%%   Pr(z) = \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{z^2}{2}\right]
+%% \end{equation*}
+
+%% And the function:
+%% \begin{equation*}
+%%   x = f[z] = \frac{1}{1 + \exp[-z]}
+%% \end{equation*}
+
+%% We can derive the inverse function $f^{-1}(x)$:
+%% \begin{equation*}
+%%   z = f^{-1}(x) = -\log \left(\frac{1}{x} - 1\right)
+%% \end{equation*}
+
+%% The formula for the change of variables in probability densities is:
+
+%% \begin{equation*}
+%%   Pr(x) = Pr(f^{-1}(x)) \left| \frac{df^{-1}(x)}{dx} \right|
+%% \end{equation*}
+
+%% By differentiating $f^{-1}(x)$ with respect to $x$, we get $\frac{df^{-1}(x)}{dx} = \frac{1}{x(1-x)}$.
+
+%% Substituting $f^{-1}(x)$ and $\frac{df^{-1}(x)}{dx}$ into the formula, we get:
+
+%% \begin{equation*}
+%%   Pr(x) = \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{(-\ln(\frac{1}{x} - 1))^2}{2}\right] \cdot \frac{1}{x(1-x)}
+%% \end{equation*}
+
+%% This is the probability density function of the transformed distribution.
+
 \subsection*{Exercise 4.4 (Normalizing Flows: Fixed point theorem)}
-Problem 16.11 from Chapter 16 of ”Understanding Deep Learning”
+
+\subsubsection*{Problem 16.11}
+
+Write out the expression for the KL divergence in the equation:
+
+\begin{equation*}
+  \phi = \argmin \left[ KL \left[ \sum_{i=1}^I \delta[x-f[z_i,\phi]] || q(x) \right] \right]
+\end{equation*}
+
+Why does it not matter if we can only evaluate the probability q(x) up to a scaling factor $\kappa$? Does the network have to be invertible to minimize this loss function? Explain your reasoning.
 
 \subsection*{Exercise 4.5 (Diffusion generative models)}
-Problems 18.1 and 18.2 from Chapter 18 of ”Understanding Deep Learning”
+
+\subsubsection*{Problem 18.1}
+
+Show that if $Var[x_{t-1}] = \textbf{I}$ and we use the update:
+
+\begin{equation*}
+  x_t = \sqrt{1-\beta_t}\cdot x_{t-1} + \sqrt{\beta_t} \cdot \epsilon_t,
+\end{equation*}
+
+then $Var[x_t] = \textbf{I}$ so the variance stays the same.
+
+\subsubsection*{Problem 18.2}
+
+Consider the variable:
+
+$$z = a \cdot \epsilon_1 + b \cdot \epsilon_2$$
+
+where both $\epsilon_1$ and $\epsilon_2$ are drawn from independent standard normal distributions with mean zero and unit variance. Show that:
+
+\begin{align*}
+  \mathbb{E}[z] &= 0 \\
+  Var[z] &= a^2 + b^2
+\end{align*}
+
+so we could equivalently compute $z = \sqrt{a^2 + b^2} \cdot \epsilon$ where $\epsilon$ is also drawn from a standard normal distribution.
 

### 2023-06-19 assignment-4/article.tex
Completed solution to 16.2
Add a solution to 16.11

@@
   p_X(x) &= p_Z(z) \left|\frac{dz}{dx}\right| \\
-  &= \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{1}{2}(\log(x(1-x)))^2 \right] \cdot \frac{1-2x}{x(1-x)} \\
-  &= \frac{1}{\sqrt{2\pi}} \exp \left[\log(x(1-x)))^{-\frac{1}{2} \cdot 2} \right] \cdot \frac{1-2x}{x(1-x)} \\
-  &= \frac{1}{\sqrt{2\pi}} \exp \left[\log\left(\frac{1}{x(1-x)}\right) \right] \cdot \frac{1-2x}{x(1-x)} \\
-  &= \frac{1}{\sqrt{2\pi}} \frac{1}{x(1-x)} \cdot \frac{1-2x}{x(1-x)} \\
-  &= \frac{1}{\sqrt{2\pi}} \frac{1-2x}{x^2(1-x)^2}
+  &= \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{1}{2}(\log(x(1-x)))^2 \right] \cdot \frac{|1-2x|}{x(1-x)}
 \end{align*}
 
-%% \subsubsection*{Problem 16.2 solution - attempt 2}
+The expression looks a bit nasty but I don't think it's possible to simplify further.
 
-%% Given the standard normal distribution:
-%% \begin{equation*}
-%%   Pr(z) = \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{z^2}{2}\right]
-%% \end{equation*}
-
-%% And the function:
-%% \begin{equation*}
-%%   x = f[z] = \frac{1}{1 + \exp[-z]}
-%% \end{equation*}
-
-%% We can derive the inverse function $f^{-1}(x)$:
-%% \begin{equation*}
-%%   z = f^{-1}(x) = -\log \left(\frac{1}{x} - 1\right)
-%% \end{equation*}
-
-%% The formula for the change of variables in probability densities is:
-
-%% \begin{equation*}
-%%   Pr(x) = Pr(f^{-1}(x)) \left| \frac{df^{-1}(x)}{dx} \right|
-%% \end{equation*}
-
-%% By differentiating $f^{-1}(x)$ with respect to $x$, we get $\frac{df^{-1}(x)}{dx} = \frac{1}{x(1-x)}$.
-
-%% Substituting $f^{-1}(x)$ and $\frac{df^{-1}(x)}{dx}$ into the formula, we get:
-
-%% \begin{equation*}
-%%   Pr(x) = \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{(-\ln(\frac{1}{x} - 1))^2}{2}\right] \cdot \frac{1}{x(1-x)}
-%% \end{equation*}
-
-%% This is the probability density function of the transformed distribution.
+\newpage
 
@@
 Why does it not matter if we can only evaluate the probability q(x) up to a scaling factor $\kappa$? Does the network have to be invertible to minimize this loss function? Explain your reasoning.
+
+\subsubsection*{Problem 16.11 solution}
+
+Writing out the KL divergence in full gives:
+
+\begin{equation*}
+  \phi = \argmin \left[ \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] \log \frac{\sum_{i=1}^I \delta[x-f[z_i,\phi]]}{q(x)} dx \right]
+\end{equation*}
+
+If we replace $q(x)$ with $\kappa q(x)$ in the above equation it can be factored out as an additive constant in the following way:
+
+\begin{align*}
+  \phi &= \argmin \left[ \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] \log \frac{\sum_{i=1}^I \delta[x-f[z_i,\phi]]}{\kappa q(x)} dx \right] \\
+  &= \argmin \left[ \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] \log \frac{\sum_{i=1}^I \delta[x-f[z_i,\phi]]}{q(x)} dx - \log \kappa \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] dx \right] \\
+  &= \argmin \left[ \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] \log \frac{\sum_{i=1}^I \delta[x-f[z_i,\phi]]}{q(x)} dx - \log \kappa \right]
+\end{align*}
+
+Where the last line holds since the probability density $\sum_{i-1}^I \delta[x-f[z_i,\phi]]$ integrates to 1 over its domain.
+Since the result of multiplying $q(x)$ by a constant $\kappa$ only adds a constant term to the KL-divergence, the $\phi$ which minimizes this divergence is the same.
+
+The network does not need to be invertible to minimize the KL divergence.
+Since we generate the samples $x_i = f[z_i , \phi]$ directly from the latent variables $z_i$, we do not need to perform an inverse operation from the data space to the latent space, which would require the network to be invertible.
 

### 2023-06-19 assignment-4/article.tex
Whitespace fixes
Solve 18.1

@@
 
+\newpage
+
 \subsection*{Exercise 4.5 (Diffusion generative models)}
@@
 then $Var[x_t] = \textbf{I}$ so the variance stays the same.
+
+\subsubsection*{Problem 18.1 solution}
+
+We proceed by calculating $Var[x_t]$ using the standard laws of variance.
+In addition, we use the fact that $x_{t-1}$ and $\epsilon_t$ are independent, and that $Var[\epsilon_t] = \textbf{I}$ since $\epsilon_t$ is drawn from a standard multivariate normal distribution.
+
+\begin{align*}
+  Var[x_t] &= Var[\sqrt{1-\beta_t}\cdot x_{t-1} + \sqrt{\beta_t} \cdot \epsilon_t] \\
+  &= Var[\sqrt{1-\beta_t}\cdot x_{t-1}] + Var[\sqrt{\beta_t} \cdot \epsilon_t] \\
+  &= (1-\beta_t) Var[x_{t-1}] + \beta_t Var[ \cdot \epsilon_t] \\
+  &= (1-\beta_t) \textbf{I} + \beta_t \textbf{I} \\
+  &= \textbf{I}
+\end{align*}
 

### 2023-06-19 assignment-4/article.tex
Remove comments
Solve 18.2

@@
 \end{itemize}
-
-%% \subsubsection*{Explicit GAN loss}
-
-%% From \citet{prince2023understanding} the GAN loss function is as follows:
-
-%% %% This is the characterisation of the saddle points
-%% \begin{equation*}
-%% \hat{\theta} = \textrm{argmax}_\theta \left[ \min_\phi \left[ \sum_j - \log(1 - D(G(z_j, \theta), \phi))) - \sum_i \log(D(x_i, \phi))) \right] \right]
-%% \end{equation*}
-
-%% where $\theta$ represents the weights of the generator function $G$, and $\phi$ is the weights of the discriminator $D$, and $\sigma$ is the sigmoid logistic function.
 
@@
 
+\subsubsection*{Problem 18.2 solution}
+
+$\mathbb{E}[z] = 0$ holds by linearity of expectation since $\mathbb{E}[\epsilon_1] = \mathbb{E}[\epsilon_2] = 0$.
+
+We calculate $Var[z]$ by hand, using the facts that $\epsilon_1$ and $\epsilon_2$ are independent, so $Cov[\epsilon_1, \epsilon_2] = 0$, and $Var[\epsilon_1] = Var[\epsilon_2] = 1$.
+
+\begin{align*}
+  Var[z] &= Var[a \epsilon_1 + b \epsilon_2] \\
+  &= a^2 Var[\epsilon_1] + b^2 Var[\epsilon_2] + 2ab Cov[\epsilon_1, \epsilon_2] \\
+  &= a^2 Var[\epsilon_1] + b^2 Var[\epsilon_2] \\
+  &= a^2 + b^2
+\end{align*}
+
 %----------------------------------------------------------------------------------------

### 2023-06-20 assignment-3/article.tex
Add assignment including solution to 3.1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Arsclassica Article
% LaTeX Template
% Version 1.1 (1/8/17)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Lorenzo Pantieri (http://www.lorenzopantieri.net) with extensive modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
10pt, % Main document font size
a4paper, % Paper type, use 'letterpaper' for US Letter paper
oneside, % One page layout (no page indentation)
%twoside, % Two page layout (page indentation for binding and different headers)
headinclude,footinclude, % Extra spacing for the header and footer
BCOR5mm, % Binding correction
]{scrartcl}

\usepackage{mdframed}
\usepackage{amsmath}
\DeclareMathOperator*{\argmax}{argmax} % thin space, limits underneath in displays
\DeclareMathOperator*{\argmin}{argmin} % thin space, limits underneath in displays
\DeclareMathOperator*{\sig}{sig}

\input{structure.tex} % Include the structure.tex file which specified the document structure and layout

\hyphenation{Fortran hy-phen-ation} % Specify custom hyphenation points in words with dashes where you would like hyphenation to occur, or alternatively, don't put any dashes in a word to stop hyphenation altogether

%----------------------------------------------------------------------------------------
%	TITLE AND AUTHOR(S)
%----------------------------------------------------------------------------------------

\title{\normalfont\spacedallcaps{Language reclamation: Literature review}} % The article title

%\subtitle{Subtitle} % Uncomment to display a subtitle

\author{\spacedlowsmallcaps{Caleb Moses*}} % The article author(s) - author affiliations need to be specified in the AUTHOR AFFILIATIONS block

\date{} % An optional date to appear under the author(s)

\newenvironment{problem}[2][Problem]
               { \begin{mdframed}[backgroundcolor=gray!20] \textbf{#1 #2} \\}
               {  \end{mdframed}}

% Define solution environment
\newenvironment{solution}
               {\textit{Solution:}}
               {}

%----------------------------------------------------------------------------------------

\title{MATH/COMP 562 ASSIGNMENT 3}
\author{Caleb Moses}

\begin{document}

\maketitle

%% \section*{Instructions}
%% Submit your solutions on MyCourses course page. You can get help from other students, but you should do the write up yourself. Code solutions can be a PDF of a workbook, you can use any programming language, but NumPy is recommended.

%% \section*{VC dimension}
%% Refer to “Foundations of Machine Learning,” by Mohri, Chapter 3.

\subsection*{Exercise 3.1 (Rademacher identities)}
\begin{problem}{Mohri Exercise 3.8. part (c) only}
Fix $m \geq 1$. Prove the following identity for any $\alpha \in \mathbb{R}$ and any two hypothesis sets $\mathcal{H}$ and $\mathcal{H}_0$ of functions mapping from $\mathcal{X}$ to $\mathbb{R}$:

$\mathfrak{R}_m(\{\max(h, h_0 ): h \in \mathcal{H}, h_0 \in \mathcal{H}_0 \}) \leq \mathfrak{R}_m (\mathcal{H}) + \mathfrak{R}_m (\mathcal{H}_0)$, where $\max(h, h_0 )$ denotes the function $x \mapsto max_{x\in \mathcal{X}} (h(x), h_0 (x))$ (Hint: you could use the identity $\max(a, b) = \frac{1}{2} [a + b + |a - b|]$ valid for all $a, b \in \mathbb{R}$ and Talagrand’s contraction lemma (see lemma 5.7)).
\end{problem}

\begin{solution}
Let $\mathcal{H}_{\max} = \{\max(h, h_0 ): h \in \mathcal{H}, h_0 \in \mathcal{H_0} \}$ be the class of functions that take the maximum of the predictions of a function $h$ from $\mathcal{H}$ and a function $h_0$ from $\mathcal{H}_0$.

Now we calculate the Rademacher complexity of $\mathcal{H}_{\max}$:

\begin{align}
%% \mathfrak{R}_m(\mathcal{H}_{\max}) &= \mathbb{E}\left[\sup_{g\in\mathcal{H}{\max}}\frac{1}{m} \sum_{i=1}^m \sigma_i g(z_i)\right] \\
\mathfrak{R}_m(\mathcal{H}_{\max}) &= \mathbb{E} \left[ \sup_{h\in\mathcal{H}, h_0\in\mathcal{H}_0} \frac{1}{m} \sum_{i=1}^m \sigma_i \max(h(z_i), h_0(z_i)) \right] \\
&= \mathbb{E} \left[ \sup_{h\in\mathcal{H}, h_0\in\mathcal{H}_0} \frac{1}{m} \sum_{i=1}^m \frac{\sigma_i}{2} (h(z_i) + h_0(z_i) + |h(z_i) - h_0(z_i)|) \right] \\
%% &\leq \mathbb{E} \left[ \sup_{h\in\mathcal{H}, h_0\in\mathcal{H}_0} \frac{1}{2m} \sum_{i=1}^m \sigma_i (h(z_i) + h_0(z_i)) + \frac{1}{2m} \sum_{i=1}^m \sigma_i |h(z_i) - h_0(z_i)| \right] \\
&\leq \frac{1}{2} \mathbb{E}\left[ \sup_{h\in\mathcal{H}}\frac{1}{m} \sum_{i=1}^m \sigma_i h(z_i)\right] + \mathbb{E}\left[\sup_{h_0\in\mathcal{H}_0}\frac{1}{m} \sum_{i=1}^m \sigma_i h_0(z_i)\right] + \\
&\qquad \frac{1}{2} \mathbb{E}\left[\sup_{h\in\mathcal{H}, h_0\in\mathcal{H}_0} \frac{1}{m} \sum_{i=1}^m \sigma_i |h(z_i) - h_0(z_i)|\right] \\
&\leq \frac{1}{2} \mathbb{E}\left[ \sup_{h\in\mathcal{H}}\frac{1}{m} \sum_{i=1}^m \sigma_i h(z_i)\right] + \frac{1}{2}\mathbb{E}\left[\sup_{h_0\in\mathcal{H}_0}\frac{1}{m} \sum_{i=1}^m \sigma_i h_0(z_i)\right] + \\
&\qquad \frac{1}{2} \mathbb{E}\left[\sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i |h(z_i)|\right] + \frac{1}{2} \mathbb{E}\left[\sup_{h_0\in\mathcal{H}_0} \frac{1}{m} \sum_{i=1}^m \sigma_i |h_0(z_i)|\right] \\
&\leq \frac{\mathfrak{R}_m(\mathcal{H})}{2} + \frac{\mathfrak{R}_m(\mathcal{H}_0)}{2} + \frac{1 \cdot \mathfrak{R}_m(\mathcal{H})}{2} + \frac{1 \cdot \mathfrak{R}_m(\mathcal{H}_0)}{2} \\
&= \mathfrak{R}_m(\mathcal{H}) + \mathfrak{R}_m(\mathcal{H}_0)
\end{align}

(1) is the definition of Rademacher complexity for $\mathcal{H}_{\max}$. We take advantage of the hint identity in (2) and then in (3) and (4) we apply the linearity of expectation and sub-additivity of the supremum. In (5) and (6) we apply the triangle inequality to the absolute difference between $h$ and $h_0$, and then in (7) we use the definition of Rademacher complexity to rewrite the first two terms, and the last two require Talagrand's contraction lemma with the fact that the Lipschitz constant of the absolute value function is 1. Simplifying gives the sum of Rademacher complexities of $\mathcal{H}$ and $\mathcal{H}_0$ as desired.

\end{solution}

\subsection*{Exercise 3.2 (VC dimension of finite hypothesis sets)}
Mohri Exercise 3.14.

\subsection*{Exercise 3.3 (VC dimension of closed balls).}
Mohri Exercise 3.17.

\subsection*{Exercise 3.4 (function class with infinite VC dimension).}
Mohri Exercise 3.20.

\section*{Stability Theory}
Refer to posted class notes and Ch 13 of Understanding Machine Learning (Shalev-Shwartz).

Exercise 3.5 (Strong Convexity). Recall that a function $f$ is $\lambda$-strongly convex if
\[
\alpha f(w) + (1 - \alpha)f(u) \geq f(\alpha w + (1 - \alpha)u) + \frac{\lambda}{2} \alpha(1 - \alpha)\|w - u\|^2, \quad \forall u, w \in \mathbb{R}^d , \alpha \in [0, 1]
\]

\begin{enumerate}
    \item Prove that $f(w) = \|w\|^2$ is $\lambda$-strongly convex with $\lambda = 2$.

    \[\] % Space to write solution

    \item Give an example of a convex function which is not $\lambda$-strongly convex for any $\lambda > 0$.

    \[\] % Space to write solution

    \item Let $f : \mathbb{R}^d \rightarrow \mathbb{R}$ be convex. Prove that $g(w, \lambda) = f(w) + \frac{\lambda}{2}\|w\|^2$ is $\lambda$-strongly convex.

    \[\] % Space to write solution

    \item Prove that $g$ is $\lambda$-strongly convex if and only if $f(w) = g(w) - \frac{\lambda}{2}\|w\|^2$ is convex.

    \[\] % Space to write solution
\end{enumerate}

% Continuing from the previous LaTeX document

\section*{Continuation of Exercises}

\begin{enumerate}
\setcounter{enumi}{5}
    \item Exercise 3.6 (Random Variables and Strong Convexity). Let $X \in \mathbb{R}$ be a discrete vector-valued random variable, with distribution
    \[
    \rho(X) =
    \begin{cases}
      w, & \text{with probability }\alpha \\
      u, & \text{with probability }(1 - \alpha)
    \end{cases}
    \]
    \begin{enumerate}
        \item Find $E[X]^2$, $E[X^2]$, $Var(X)$. Show that the expression $E[X^2] - (E[X])^2 = Var(X)$ corresponds to
        \[
        \alpha f(w) + (1 - \alpha) f(u) - f(\alpha w + (1 - \alpha) u) = \alpha (1 - \alpha) \|w - u\|^2
        \]
        with $f(w) = w^2$.

        \[\] % Space to write solution

        \item Suppose $f$ is $\lambda$ strongly convex. Let $X$ be as above. Show that
        \[
        E[f(X)] - f(E[X]) \geq \frac{\lambda}{2} Var(X)
        \]

        \[\] % Space to write solution
    \end{enumerate}

    \item Exercise 3.7 (Algorithmic Stability for Ridge Regression updated for clarity). Write,
    \[
    f(w, S) = L_S(w) = \frac{1}{m} \sum_{i=1}^{m} (w \cdot x_i - y_i)^2
    \]
    Consider the datasets $S_1$, $S_2$, which differ by exactly one data point. Write
    \[
    z_1 = (x_1, y_1) \in S_1, \quad z_2 = (x_2, y_2) \in S_2
    \]
    for the different points. Define $g(w, S) = f(w, S) + \frac{\lambda}{2} \|w\|^2$ and
    \[
    w_i = A(S_i) = \arg\min_w g(w, S_i), \quad i = 1, 2
    \]
    \begin{enumerate}
        \item Explain why $g(w, S)$ is $\lambda$-strongly convex (as a function of $w$).

        \[\] % Space to write solution

        \item Explain how strong convexity is applied to obtain
        \[
        g(w_2, S_1) - g(w_1, S_1) \geq \frac{\lambda}{2} \|w_1 - w_2\|^2
        \]
        \[
        g(w_1, S_2) - g(w_2, S_2) \geq \frac{\lambda}{2} \|w_1 - w_2\|^2
        \]

        \[\] % Space to write solution

        \item Simplify to obtain
        \[
        \lambda \|w_1 - w_2\|^2 \leq \frac{1}{m} | \ell(w_2, z_1) - \ell(w_1, z_1) + \ell(w_1, z_2) - \ell(w_2, z_2) |
        \]
        for $\ell(w, z) = (w \cdot x - y)^2$.

        \[\] % Space to write solution

        \item Assume that $\|x\| \leq C_x$ and $|y| \leq C_y$ for all data $x$, $y$, and that $\|w_1\|, \|w_2\| \leq C_w$. Bound the RHS of the previous inequality. (Hint: apply the identity $(a^2 - b^2) = (a - b)(a + b)$ to the loss.)

        \[\] % Space to write solution

        \item Conclude that the Ridge Regression problem is replace one stable (ROS) and give the rate as a function of $\lambda$, $m$.

        \[\] % Space to write solution
    \end{enumerate}

    \item Exercise 3.8 (Coding exercises). Using your choice of dataset (one dimensional is fine), perform Ridge regression. Find $w = A(S)$. Change one data point to obtain $S'$. Compute $w' = A(S')$. Changing the data by different amounts, check the stability bounds. Plot the $\|w - w'\|$ against the bound.

    \[\] % Space to write solution
\end{enumerate}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\bibliography{sample.bib} % The file containing the bibliography

\end{document}

### 2023-06-20 assignment-3/article.tex
Remove comments - clean up the proof
Update 3.1 explanation

@@
 \begin{align}
-%% \mathfrak{R}_m(\mathcal{H}_{\max}) &= \mathbb{E}\left[\sup_{g\in\mathcal{H}{\max}}\frac{1}{m} \sum_{i=1}^m \sigma_i g(z_i)\right] \\
 \mathfrak{R}_m(\mathcal{H}_{\max}) &= \mathbb{E} \left[ \sup_{h\in\mathcal{H}, h_0\in\mathcal{H}_0} \frac{1}{m} \sum_{i=1}^m \sigma_i \max(h(z_i), h_0(z_i)) \right] \\
 &= \mathbb{E} \left[ \sup_{h\in\mathcal{H}, h_0\in\mathcal{H}_0} \frac{1}{m} \sum_{i=1}^m \frac{\sigma_i}{2} (h(z_i) + h_0(z_i) + |h(z_i) - h_0(z_i)|) \right] \\
-%% &\leq \mathbb{E} \left[ \sup_{h\in\mathcal{H}, h_0\in\mathcal{H}_0} \frac{1}{2m} \sum_{i=1}^m \sigma_i (h(z_i) + h_0(z_i)) + \frac{1}{2m} \sum_{i=1}^m \sigma_i |h(z_i) - h_0(z_i)| \right] \\
 &\leq \frac{1}{2} \mathbb{E}\left[ \sup_{h\in\mathcal{H}}\frac{1}{m} \sum_{i=1}^m \sigma_i h(z_i)\right] + \mathbb{E}\left[\sup_{h_0\in\mathcal{H}_0}\frac{1}{m} \sum_{i=1}^m \sigma_i h_0(z_i)\right] + \\
@@
 We take advantage of the hint identity in (2) and then in (3) and (4) we apply the linearity of expectation and sub-additivity of the supremum.
-In (5) and (6) we apply the triangle inequality to the absolute difference between $h$ and $h_0$, and then in (7) we use the definition of Rademacher complexity to rewrite the first two terms, and the last two require Talagrand's contraction lemma with the fact that the Lipschitz constant of the absolute value function is 1.
+In (6) we apply the triangle inequality to the absolute difference between $h$ and $h_0$, and use sub-additivity of the supremum and linearity of expectation again to split it into two expectations.
+Then in (7) we use the definition of Rademacher complexity to rewrite the first two terms, and the last two require Talagrand's contraction lemma with the fact that the Lipschitz constant of the absolute value function is 1.
 Simplifying gives the sum of Rademacher complexities of $\mathcal{H}$ and $\mathcal{H}_0$ as desired.

### 2023-06-20 assignment-3/article.tex
Remove old title
No need to say "Problem" for every problem
Add problem statements for VC dimension

@@
 
-\title{\normalfont\spacedallcaps{Language reclamation: Literature review}} % The article title
-
 %\subtitle{Subtitle} % Uncomment to display a subtitle
@@
 
-\newenvironment{problem}[2][Problem]
+\newenvironment{problem}[2][]
                { \begin{mdframed}[backgroundcolor=gray!20] \textbf{#1 #2} \\}
@@
 \subsection*{Exercise 3.2 (VC dimension of finite hypothesis sets)}
-Mohri Exercise 3.14.
+\begin{problem}{Mohri Exercise 3.14}
+VC-dimension of finite hypothesis sets. Show that the VC-dimension of a finite hypothesis set $\mathcal{H}$ is at most $\log_2|\mathcal{H}|$.
+\end{problem}
+
+\begin{solution}
+Here is a solution
+\end{solution}
 
 \subsection*{Exercise 3.3 (VC dimension of closed balls).}
-Mohri Exercise 3.17.
+\begin{problem}{Mohri Exercise 3.17}
+Show that the VC-dimension of the set of all closed balls in $\mathbb{R}_n$, i.e.
+sets of the form ${x \in \mathbb{R}_n : kx - x0 k2 \leq r}$ for some $x_0 \in \mathbb{R}_n$ and $r \geq 0$, is less than or equal to $n + 2$.
+\end{problem}
 
-\subsection*{Exercise 3.4 (function class with infinite VC dimension).}
-Mohri Exercise 3.20.
+\begin{solution}
+Here is a solution
+\end{solution}
+
+\subsection*{Exercise 3.4 (Function class with infinite VC dimension).}
+\begin{problem}{Mohri Exercise 3.20}
+Consider the hypothesis family of sine functions: $\{x \to sin(\omega x): \omega \in \mathbb{R}\}$.
+\begin{enumerate}[label=(\alph*)]
+\item Show that for any $x \in \mathbb{R}$ the points $x$, $2x$, $3x$ and $4x$ cannot be shattered by this family of sine functions.
+\item Show that the VC-dimension of the family of sine functions is infinite. (Hint: show that $\{2-i: i \leq m\}$ can be shattered for any $m > 0$.)
+\end{enumerate}
+\end{problem}
+
+\begin{solution}
+Here is a solution
+\end{solution}
 

### 2023-06-23 assignment-3/article.tex
Solve VC dimension of closed balls

@@
 \begin{solution}
-Here is a solution
+  Let $\mathcal{H}$ be a finite hypothesis set that shatters an instance space $X$.
+Then each element of $\mathcal{H}$ corresponds to a possible labelling of $X$ and there are $2^{|X|}$ such labellings.
+Since $\mathcal{H}$ shatters $X$ we require $|\mathcal{H}| \geq 2^{|X|}$, since we need at least as many hypotheses as there are labellings for $X$.
+Equivalently, $\log_2{|\mathcal{H}|} \geq |X|$ so the largest set $X$ that $\mathcal{H}$ can shatter has size $\log_2 |\mathcal{H}|$ which gives the desired upper bound on the VC-dimension.
 \end{solution}
@@
 Show that the VC-dimension of the set of all closed balls in $\mathbb{R}_n$, i.e.
-sets of the form ${x \in \mathbb{R}_n : kx - x0 k2 \leq r}$ for some $x_0 \in \mathbb{R}_n$ and $r \geq 0$, is less than or equal to $n + 2$.
+sets of the form ${x \in \mathbb{R}_n : ||x-x_0|| \leq r}$ for some $x_0 \in \mathbb{R}_n$ and $r \geq 0$, is less than or equal to $n + 2$.
 \end{problem}
@@
 \begin{solution}
-Here is a solution
+Fix $r>0$, then the constraint $||x-a||^2<r$ is equivalent to:
+\begin{align*}
+  ||x-a||^2 - r &< 0 \\
+  (x-a)^T(x-a) - r &< 0 \\
+  ||x||^2 - 2(x\cdot a) + ||a||^2 - r &< 0 \\
+  \sum_{i=1}^n x_i^2 - 2 \sum_{i=1}^n a_i x_i + \sum_{i=1}^n a_i^2 - r &< 0
+\end{align*}
 \end{solution}
+
+We will show that there are enough free variables to capture $n+1$ arbitrary points. The last line is equivalent to $WX - B < 0$ with:
+
+\begin{align*}
+  W &= \begin{bmatrix}
+    1 \\
+    -2a_1 \\
+    ... \\
+    -2a_n
+   \end{bmatrix} & X = \begin{bmatrix}
+     \sum_{i=1}^n x_i^2 \\
+     x_1 \\
+     ... \\
+     x_n
+   \end{bmatrix} &&  B = \sum_{i=1}^n r - a_i^2
+\end{align*}
+
+There are n+1 linearly independent equations with n+1 unknowns (1 for each $a_i$ from 1 to n, plus $r$).
+So the VC dimension of closed balls in $\mathbb{R}^n$ is at most equal to the VC dimension of hyperplanes in $\mathbb{R}^{n+1}$, which is $n+2$.
 

### 2023-06-23 assignment-3/article.tex
Solve VC dimension of sine curves

@@
 \begin{problem}{Mohri Exercise 3.20}
-Consider the hypothesis family of sine functions: $\{x \to sin(\omega x): \omega \in \mathbb{R}\}$.
+Consider the hypothesis family of sine functions: $\mathcal{H} = \{x \to sin(\omega x): \omega \in \mathbb{R}\}$.
 \begin{enumerate}[label=(\alph*)]
 \item Show that for any $x \in \mathbb{R}$ the points $x$, $2x$, $3x$ and $4x$ cannot be shattered by this family of sine functions.
-\item Show that the VC-dimension of the family of sine functions is infinite. (Hint: show that $\{2-i: i \leq m\}$ can be shattered for any $m > 0$.)
+\item Show that the VC-dimension of the family of sine functions is infinite. (Hint: show that $\{2^{-i}: i \leq m\}$ can be shattered for any $m > 0$.)
 \end{enumerate}
@@
 \begin{solution}
-Here is a solution
+\begin{enumerate}[label=(\alph*)]
+  \item Fix $x\in\mathbb{R}$.
+Consider the dichotomy $\{(x, -), (2x, -), (3x, +), (4x, -)\}$.
+Assume towards contradiction that $\mathcal{H}$ realises this dichotomy, so there exists $\omega\in\mathbb{R}$ satisfying $\sin(\omega x) < 0$, $\sin(2 \omega x) < 0$, $\sin(3 \omega x) > 0$ and $\sin(4 \omega x) < 0$.
+
+  Using the double angle formulae for sin and cos gives:
+  \begin{align*}
+    \sin(4\omega x) &= 2 \sin(2\omega x)\cos(2\omega x) \\
+    &= 2\sin(2\omega x) (1 - 2\sin^2(\omega x))
+  \end{align*}
+
+  Since $\sin(4\omega x) < 0$ and $\sin(2\omega x) < 0$ by assumption, $1 - 2\sin^2(\omega x) > 0$ must hold otherwise the right hand side is positive while the left hand side is negative.
+Therefore $\sin^2(\omega x) < \frac{1}{2}$.
+
+  Next we use the triple angle formula for $\sin$ to calculate the following:
+  \begin{align*}
+    \sin(3\omega x) &= \sin(\omega x)(3 - 4\sin^2(\omega x))
+  \end{align*}
+
+  Since $\sin(3\omega x) > 0$ and $\sin(\omega x) < 0$, we require $3 - 4\sin^2(\omega x) < 0$ otherwise the left hand side is positive and the right hand side is negative.
+Therefore $\sin^2(\omega x) > \frac{3}{4}$.
+This contradicts $\sin^2(\omega x) < \frac{1}{2}$, so there is no $\omega$ such that $\mathcal{H}$ shatters $x$, $2x$, $3x$ and $4x$.
+
+\item Let $m \in \mathbb{N}$ and consider the points $\{2^{-i}: i \leq m\}$ with arbitrary labels $(y_1, ...
+y_{m}) \in \{-1, +1\}^m$.
+Let $y_i^\prime = \frac{1-y_i}{2}$ and choose $\omega = \pi(1 + \sum_{i=1}^m 2^{i}y_i^\prime)$.
+We will show that this is sufficient to correctly classify our chosen labels.
+then for any $j\in[1,m]$ we have the following:
+    \begin{align*}
+      \omega 2^{-j} &= 2^{-j} \pi\left[1 + \sum_{i=1}^m 2^{i}y_i^\prime\right] \\
+      &= \pi \left[2^{-j} + \sum_{i=1}^m 2^{i-j}y_i^\prime\right] \\
+      &= \pi \left[2^{-j} + y_j^\prime + \sum_{i=1}^{j-1} 2^{i-j}y_i^\prime + \sum_{i=j+1}^m 2^{i-j}y_i^\prime \right] \\
+      &= \pi \left[2^{-j} + y_j^\prime + \sum_{i=1}^{j-1} 2^{-i}y_{j-i}^\prime + \sum_{i=1}^{m-j} 2^iy_{i+j}^\prime \right] \\
+      &= \pi \left[2^{-j} + y_j^\prime + \sum_{i=1}^{j-1} 2^{-i}y_{j-i}^\prime\right] + 2\pi \sum_{i=1}^{m-j} 2^{i-1}y_{i+j}^\prime
+    \end{align*}
+  The last term contributes multiples of $2\pi$ to $\omega 2^{-i}$ and so can be ignored. Since $y_i\in\{0,1\}$, the remaining term can be bound in the following way:
+
+  \begin{align*}
+    \pi \left[2^{-j} + y_j^\prime + \sum_{i=1}^{j-1} 2^{-i}y_{j-i}^\prime\right] &\leq \pi \left[ y_j^\prime + \sum_{i=1}^{j} 2^{-i}\right] < \pi(1 + y_j^\prime) \\
+    \pi \left[2^{-j} + y_j^\prime + \sum_{i=1}^{j-1} 2^{-i}y_{j-i}^\prime\right] &> \pi y_j^\prime
+  \end{align*}
+
+  Now if $y_j = 1$ then $y_j^\prime = 0$ and $0 < \omega 2^{-j} < \pi$ which implies $\sin(\omega 2^{-j}) > 0$.
+Similarly, if $y_j = -1$, then $y_j^\prime = 1$, $\pi < \omega 2^{-j} < 2\pi$ and we have that $\sin(\omega 2^{-j}) < 0$ as desired.
+
+\end{enumerate}
 \end{solution}
 
+\newpage
+
 \section*{Stability Theory}
-Refer to posted class notes and Ch 13 of Understanding Machine Learning (Shalev-Shwartz).
 

### 2023-06-23 assignment-3/article.tex
Fix section headers
Tidy up exercise 3.5

@@
 
-%% \section*{Instructions}
-%% Submit your solutions on MyCourses course page.
-You can get help from other students, but you should do the write up yourself.
-Code solutions can be a PDF of a workbook, you can use any programming language, but NumPy is recommended.
-
-%% \section*{VC dimension}
-%% Refer to “Foundations of Machine Learning,” by Mohri, Chapter 3.
-
+\section*{Rademacher Complexity}
 \subsection*{Exercise 3.1 (Rademacher identities)}
@@
 \end{solution}
+
+\section{VC dimension}
 
@@
 \section*{Stability Theory}
-
-Exercise 3.5 (Strong Convexity). Recall that a function $f$ is $\lambda$-strongly convex if
-\[
-\alpha f(w) + (1 - \alpha)f(u) \geq f(\alpha w + (1 - \alpha)u) + \frac{\lambda}{2} \alpha(1 - \alpha)\|w - u\|^2, \quad \forall u, w \in \mathbb{R}^d , \alpha \in [0, 1]
-\]
-
-\begin{enumerate}
+Recall that a function $f$ is $\lambda$-strongly convex if for all $u, w \in \mathbb{R}^d , \alpha \in [0, 1]$:
+$$f(\alpha w + (1 - \alpha)u) + \frac{\lambda}{2} \alpha(1 - \alpha)\|w - u\|^2 \leq \alpha f(w) + (1 - \alpha)f(u)$$
+\begin{problem}{Exercise 3.5 (Strong convexity)}
+  \begin{enumerate}[label=(\alph*)]
     \item Prove that $f(w) = \|w\|^2$ is $\lambda$-strongly convex with $\lambda = 2$.
-
-    \[\] % Space to write solution
-
     \item Give an example of a convex function which is not $\lambda$-strongly convex for any $\lambda > 0$.
-
-    \[\] % Space to write solution
-
     \item Let $f : \mathbb{R}^d \rightarrow \mathbb{R}$ be convex. Prove that $g(w, \lambda) = f(w) + \frac{\lambda}{2}\|w\|^2$ is $\lambda$-strongly convex.
-
-    \[\] % Space to write solution
-
     \item Prove that $g$ is $\lambda$-strongly convex if and only if $f(w) = g(w) - \frac{\lambda}{2}\|w\|^2$ is convex.
-
-    \[\] % Space to write solution
   \end{enumerate}
-
-% Continuing from the previous LaTeX document
-
-\section*{Continuation of Exercises}
+\end{problem}
 

### 2023-06-23 assignment-3/article.tex
Solve strong convexity a-c

@@
 
+\begin{solution}
+  \begin{enumerate}[label=(\alph*)]
+      \item We begin by calculating the left hand side of the strong Jensen's inequality:
+        \begin{align*}
+        &\hphantom{=} ||\alpha w + (1-\alpha)u)||^2 + \alpha(1-\alpha)||w-u||^2 \\
+        &= (\alpha w + (1-\alpha)u)^T(\alpha w + (1-\alpha)u) + \alpha(1-\alpha)(w-u)^T(w-u) \\
+        &= \alpha^2||w||^2 + 2\alpha(1-\alpha)(u\cdot w) + (1-\alpha)^2||u||^2 + \alpha(1-\alpha)(||w||^2 + 2(u\cdot w) + ||u||^2) \\
+        &= (\alpha^2+\alpha(1-\alpha))||w||^2 + 4\alpha(1-\alpha)(u\cdot w) + ((1-\alpha)^2 + \alpha(1-\alpha))||u||^2 \\
+        &\leq \alpha||w||^2 + 4\alpha(1-\alpha)||u||\cdot ||w|| + (1-\alpha)||u||^2 \\
+        &\leq \alpha||w||^2 + (1-\alpha)||u||^2
+        \end{align*}
+        as required.
+      \item I claim that $f(w)=|w|$ is convex, but not $\lambda$-strongly convex for any $\lambda$. By the triangle inequality, the following inequality is tight:
+        \begin{align*}
+          0 &\leq \alpha|w| + (1-\alpha)|u| -|\alpha w + (1-\alpha)u)|
+        \end{align*}
+        Strong convexity requires we can bound this term below by something quadratic in $||w-u||$ but this is not possible, so $|w|$ is not strongly convex.
+      \item We calculate Jensen's inequality for $g$: \begin{align*}
+        &\hphantom{=} g(\alpha w + (1-\alpha)u, \lambda) + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
+        &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}||\alpha w + (1-\alpha) u||^2 + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
+        &\leq f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}(\alpha||w||^2 + (1-\alpha)||u||^2 - \alpha(1-\alpha)||w-u||^2) + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
+        &\leq f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}(\alpha||w||^2 + (1-\alpha)||u||^2) \\
+        &\leq \alpha (f(w) + \frac{\lambda}{2}||w||^2) + (1-\alpha)(f(u) + \frac{\lambda}{2}||u||^2)
+      \end{align*}
+        as required.
+      \item Suppose $f(w) = g(w) - \frac{\lambda}{2}||w||^2$ is convex. We will show that $g$ is $\lambda$-strongly convex:
+        \begin{align*}
+          &\hphantom{=} g(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
+          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}||\alpha w + (1-\alpha) u||^2 + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2
+        \end{align*}
+  \end{enumerate}
+\end{solution}
+
 \begin{enumerate}
@@
 
-    \item Exercise 3.8 (Coding exercises).
-Using your choice of dataset (one dimensional is fine), perform Ridge regression.
-Find $w = A(S)$.
-Change one data point to obtain $S'$.
-Compute $w' = A(S')$.
-Changing the data by different amounts, check the stability bounds.
-Plot the $\|w - w'\|$ against the bound.
-
     \[\] % Space to write solution

### 2023-06-24 assignment-3/article.tex
Add nested enumeration
New page split
Solve 3.6
Write up random variable question

@@
 ]{scrartcl}
-
+\renewcommand{\labelenumii}{\theenumii}
+\renewcommand{\theenumii}{\theenumi.\arabic{enumii}.}
 \usepackage{mdframed}
@@
         as required.
+        \newpage
       \item Suppose $f(w) = g(w) - \frac{\lambda}{2}||w||^2$ is convex. We will show that $g$ is $\lambda$-strongly convex:
@@
           &\hphantom{=} g(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
-          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}||\alpha w + (1-\alpha) u||^2 + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2
+          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}||\alpha w + (1-\alpha) u||^2 + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
+          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}(\alpha w + (1-\alpha) u)^T(\alpha w + (1-\alpha) u) + \\
+          &\hphantom{=} \frac{\lambda}{2}\alpha(1-\alpha)(w-u)^T(w-u) \\
+          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}\left(\alpha^2||w||^2+2\alpha(1-\alpha)(u\cdot w) + (1-\alpha)^2||u||^2\right) + \\
+          &\hphantom{=} \frac{\lambda}{2}\alpha(1-\alpha)\left(||w||^2 - 2(u\cdot w) + ||u||^2\right) \\
+          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}\left(\alpha^2||w||^2 + (1-\alpha)^2||u||^2\right) + \frac{\lambda}{2}\alpha(1-\alpha)\left(||w||^2 + ||u||^2\right) \\
+          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}||w||^2\left(\alpha^2 + \alpha(1-\alpha)\right) + \frac{\lambda}{2}||u||^2\left((1-\alpha)^2 + \alpha(1-\alpha)\right) \\
+          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda\alpha}{2}||w||^2 + \frac{\lambda(1-\alpha)}{2}||u||^2 \\
+          &\leq \alpha f(w) + (1-\alpha)f(u) + \frac{\lambda\alpha}{2}||w||^2 + \frac{\lambda(1-\alpha)}{2}||u||^2 \\
+          &\leq \alpha (f(w) + \frac{\lambda}{2}||w||^2) + (1-\alpha)(f(u) + \frac{\lambda}{2}||u||^2) \\
+          &= \alpha g(w) + (1-\alpha)g(u)
         \end{align*}
+        as desired.
+
+        Now we must prove the converse. To that aim, assume $g$ is $\lambda$-strongly convex. We will now prove that $f$ is convex.
+        \begin{align*}
+          &\hphantom{=} f(\alpha w + (1-\alpha)u) \\
+          &= g(\alpha w + (1-\alpha)u) - \frac{\lambda}{2}||\alpha w + (1-\alpha)u||^2 \\
+          &\leq \alpha g(w) + (1-\alpha)g(u) - \frac{\lambda}{2}||\alpha w + (1-\alpha) u||^2 - \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
+          &\leq \alpha g(w) + (1-\alpha)g(u) - \frac{\lambda}{2}(\alpha w + (1-\alpha) u)^T(\alpha w + (1-\alpha) u) - \\
+          &\hphantom{\leq} \frac{\lambda}{2}\alpha(1-\alpha)(w-u)^T(w-u) \\
+          &\leq \alpha g(w) + (1-\alpha)g(u) - \frac{\lambda}{2}\left(\alpha^2 ||w||^2 + 2\alpha(1-\alpha)(u \cdot w) + (1-\alpha) ||u||^2\right) - \\
+          &\hphantom{\leq} \frac{\lambda}{2}\alpha(1-\alpha)(||w||^2 - 2(u\cdot w) + ||u||^2) \\
+          &\leq \alpha g(w) + (1-\alpha)g(u) - \frac{\lambda}{2}\left(\alpha^2 ||w||^2 + (1-\alpha) ||u||^2\right) - \frac{\lambda}{2}\alpha(1-\alpha)(||w||^2 + ||u||^2) \\
+          &\leq \alpha g(w) - \frac{\lambda}{2}||w||^2\left(\alpha^2 + \alpha(1-\alpha) \right)  + (1-\alpha)g(u) - \frac{\lambda}{2}||u||^2\left((1-\alpha)^2 + \alpha(1-\alpha)\right) \\
+          &\leq \alpha g(w) - \frac{\lambda\alpha}{2}||w||^2  + (1-\alpha)g(u) - \frac{\lambda(1-\alpha)}{2}||u||^2 \\
+          &\leq \alpha \left(g(w) - \frac{\lambda}{2}||w||^2\right) + (1-\alpha)\left(g(u) - \frac{\lambda}{2}||u||^2\right) \\
+          &= \alpha f(w) + (1-\alpha)f(u)
+        \end{align*}
+        so $f$ is convex.
   \end{enumerate}
@@
 
-\begin{enumerate}
-\setcounter{enumi}{5}
-    \item Exercise 3.6 (Random Variables and Strong Convexity). Let $X \in \mathbb{R}$ be a discrete vector-valued random variable, with distribution
+\begin{problem}{Exercise 3.6 (Random Variables and Strong Convexity)}
+  Let $X \in \mathbb{R}$ be a discrete vector-valued random variable, with distribution
   \[
@@
 
-        \[\] % Space to write solution
-
-        \item Suppose $f$ is $\lambda$ strongly convex. Let $X$ be as above. Show that
+      \item Suppose $f$ is $\lambda$-strongly convex. Let $X$ be as above. Show that
       \[
@@
       \]
-
-        \[\] % Space to write solution
   \end{enumerate}
+\end{problem}
 

### 2023-06-24 assignment-3/article.tex
Solve random variable questions
Write up algorithmic stability

@@
 
+\begin{solution}
+  \begin{enumerate}
+    \item
+      \begin{enumerate}
+        \item $E[X] = \alpha w + (1-\alpha)u$ therefore $E[X]^2 = ||\alpha w + (1-\alpha)u||^2$
+        \item $E[X^2] = \alpha||w||^2 + (1-\alpha) ||u||^2$
+        \item $\begin{aligned}[t]
+          Var[X] &= E[X^2] - E[X]^2 \\
+          &= \alpha||w||^2 + (1-\alpha) ||u||^2 - ||\alpha w + (1-\alpha)u||^2 \\
+          &= \alpha||w||^2 + (1-\alpha) ||u||^2 - (\alpha w + (1-\alpha)u)^T(\alpha w + (1-\alpha)u) \\
+          &= \alpha||w||^2 + (1-\alpha) ||u||^2 - \left(\alpha^2||w||^2 -2\alpha(1-\alpha)(u\cdot w) + (1-\alpha)^2||u||^2\right) \\
+          &= ||w||^2(\alpha -\alpha^2)+2\alpha(1-\alpha)(u\cdot w) + ||u||^2(1-\alpha-(1-\alpha)^2) \\
+          &= ||w||^2\alpha(1 -\alpha)+2\alpha(1-\alpha)(u\cdot w) + ||u||^2\alpha(1 - \alpha) \\
+          &= \alpha(1-\alpha)\left(||w||^2 - 2(u\cdot w) + ||u||^2\right) \\
+          &= \alpha(1-\alpha)||w-u||^2
+          \end{aligned}$
+      \end{enumerate}
+    \item $\begin{aligned}[t]
+      E[f(X)] - f(E[X]) &= \alpha f(w) + (1-\alpha)f(u) - f(\alpha w + (1-\alpha)u) \\
+      &\geq \alpha f(w) + (1-\alpha)f(u) - \alpha f(w) -(1-\alpha)f(u) + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
+      &\geq \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 = \frac{\lambda}{2}Var[x]
+      \end{aligned}$
+  \end{enumerate}
+\end{solution}
+
     \item Exercise 3.7 (Algorithmic Stability for Ridge Regression updated for clarity). Write,
@@
 
-        \[\] % Space to write solution
-
         \item Simplify to obtain
@@
 
-        \[\] % Space to write solution
-
         \item Assume that $\|x\| \leq C_x$ and $|y| \leq C_y$ for all data $x$, $y$, and that $\|w_1\|, \|w_2\| \leq C_w$.
@@
 
-        \[\] % Space to write solution
+        \item Conclude that the Ridge Regression problem is replace one stable (ROS) and give the rate as a function of $\lambda$, $m$.
+    \end{enumerate}
+\end{problem}
 
-        \item Conclude that the Ridge Regression problem is replace one stable (ROS) and give the rate as a function of $\lambda$, $m$.
+\begin{solution}
+  \begin{enumerate}
+    \item $g$ is $\lambda$-strongly convex by the proof in exercise 3.5(c)
+    \item Since $g$ is $\lambda$-strongly convex, we apply the inequality setting $\alpha=0.5$): $$g(\frac{1}{2} w_1 + \frac{1}{2}w_2, S_1) + \frac{\lambda}{4}||w_1-w_2||^2 \leq \frac{1}{2} g(w_1, S_1) + \frac{1}{2}g(w_2, S_1)$$
 
-        \[\] % Space to write solution
+      Because $w_1$ minimises $g$ at $S_1$, we have that $g(w_1, S_1) \leq g(\frac{1}{2} w_1 + \frac{1}{2} w_2, S_1)$.
+So we can substitute $g(w_1, S_1)$ in the left hand side in the following way: $$g(w_1, S_1) + \frac{\lambda}{4}||w_1-w_2||^2 \leq \frac{1}{2}g(w_1, s_1) + \frac{1}{2}g(w_2, S_1)$$
+
+      Subtracting $g(w_1, S_1)$ from both sides of the inequality, and then multiplying both sides by 2 gives the first desired inequality:
+      $$\frac{\lambda}{2}||w_1-w_2||^2 \leq g(w_2,S_1)-g(w_1, S_1)$$
+      We can obtain the other inequality by the same argument, or via symmetry.
+      $$\frac{\lambda}{2}||w_1-w_2||^2 \leq g(w_1,S_2)-g(w_2, S_2)$$
+    \item We start by rewriting the inequalities from the preceding exercise.
+      \begin{align*}
+        \frac{\lambda}{2}||w_1-w_2||^2 \leq g(w_2, z_1)-g(w_1, z_1) \\
+        \frac{\lambda}{2}||w_1-w_2||^2 \leq g(w_1, z_2)-g(w_2, z_2)
+      \end{align*}
+      Then we add both inequalities together to obtain:
+      \begin{align*}
+        \lambda ||w_1-w_2||^2 &\leq (g(w_2,z_1)-g(w_1,z_1)+g(w_1,z_2)-g(w_2,z_2)) \\
+        \hphantom{\lambda ||w_1-w_2||^2} &\leq (f(w_2,z_1) + ||w_2||^2 -f(w_1,z_1) - ||w_2||^2 \\
+        \hphantom{\lambda ||w_1-w_2||^2}&+f(w_1,z_2) + ||w_1||^2-f(w_2,z_2) - ||w_1||^2 \\
+        &\leq f(w_2,z_1)-f(w_1,z_1)+f(w_1,z_2)-f(w_2,z_2) \\
+        &\leq \frac{1}{m} (\ell(w_2,z_1)-\ell(w_1,z_1)+\ell(w_1,z_2)-\ell(w_2,z_2))
+      \end{align*}
+      where the last inequality comes from the fact that $f(w_i,S_i) = \frac{1}{m}\ell(w_i, z_i)$.
+      \item We begin by calculating an upper bound for $\ell(w,z)$: \begin{align*}
+        \ell(w, z) &= (w \cdot x - y)^2 \\
+        &= ||w\cdot x||^2 - 2 (w \cdot x) \cdot y + ||y||^2 \\
+        &\leq ||w||^2 ||x||^2 - 2||w|| ||x|| ||y|| + ||y||^2 \\
+        &\leq ||w||^2 ||x||^2 + ||y||^2 \\
+        &\leq C_w^2C_x^2 + C_y^2
+        \end{align*}
+
+        Now we use the hint:
+        \begin{align*}
+          \ell(w_2,z_1) - \ell(w_1,z_1) &= (w_2\cdot x - y)^2 - (w_1\cdot x - y)^2 \\
+          \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &= (w_2\cdot x - y - w_1\cdot x + y)(w_2\cdot x - y + w_1\cdot x - y) \\
+          \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &= ((w_2 - w_1)\cdot x)((w_1 + w_2)\cdot x - 2y)
+        \end{align*}
   \end{enumerate}
-
-    \[\] % Space to write solution
-\end{enumerate}
+\end{solution}
 

### 2023-06-25 assignment-3/article.tex
Tidy up
Fix new pages
Remove the bibliography
Fix old solutions

@@
 So the VC dimension of closed balls in $\mathbb{R}^n$ is at most equal to the VC dimension of hyperplanes in $\mathbb{R}^{n+1}$, which is $n+2$.
-
+\newpage
 \subsection*{Exercise 3.4 (Function class with infinite VC dimension).}
@@
 
-\newpage
-
 \section*{Stability Theory}
@@
   \begin{enumerate}[label=(\alph*)]
-      \item We begin by calculating the left hand side of the strong Jensen's inequality:
+      \item We begin by calculating the left hand side of the strong convexity inequality:
         \begin{align*}
@@
         &= (\alpha w + (1-\alpha)u)^T(\alpha w + (1-\alpha)u) + \alpha(1-\alpha)(w-u)^T(w-u) \\
-        &= \alpha^2||w||^2 + 2\alpha(1-\alpha)(u\cdot w) + (1-\alpha)^2||u||^2 + \alpha(1-\alpha)(||w||^2 + 2(u\cdot w) + ||u||^2) \\
-        &= (\alpha^2+\alpha(1-\alpha))||w||^2 + 4\alpha(1-\alpha)(u\cdot w) + ((1-\alpha)^2 + \alpha(1-\alpha))||u||^2 \\
-        &\leq \alpha||w||^2 + 4\alpha(1-\alpha)||u||\cdot ||w|| + (1-\alpha)||u||^2 \\
-        &\leq \alpha||w||^2 + (1-\alpha)||u||^2
+        &= \alpha^2||w||^2 + 2\alpha(1-\alpha)(u\cdot w) + (1-\alpha)^2||u||^2 + \alpha(1-\alpha)(||w||^2 - 2(u\cdot w) + ||u||^2) \\
+        &= (\alpha^2+\alpha(1-\alpha))||w||^2 + ((1-\alpha)^2 + \alpha(1-\alpha))||u||^2 \\
+        &= \alpha||w||^2 + (1-\alpha)||u||^2
         \end{align*}
-        as required.
-      \item I claim that $f(w)=|w|$ is convex, but not $\lambda$-strongly convex for any $\lambda$. By the triangle inequality, the following inequality is tight:
+      \item I claim that $f(x) = x^4$ is convex, but not $\lambda$-strongly convex for any $\lambda>0$. We know $f$ is convex since $f^{\prime\prime}(x) = 12x^2 > 0$ for all x.
+        Now to show $f$ is not $\lambda$-strongly convex, choose $x_0=0$, $x_1=w$ and $\alpha=\frac{1}{2}$.
+Suppose towards contradiction that $f$ is $\lambda$-strongly convex for some $\lambda > 0$.
+Then the following must hold for all $w\in\mathbb{R}$:
         \begin{align*}
-          0 &\leq \alpha|w| + (1-\alpha)|u| -|\alpha w + (1-\alpha)u)|
+          \left(\frac{w}{2}\right)^4 + \frac{\lambda}{8}w^2 &\leq \frac{1}{2}w^4 \\
+          \lambda &\leq \frac{8}{w^2}\left(\frac{w^4}{2} - \frac{w^4}{16}\right) = 3.5w^2
         \end{align*}
-        Strong convexity requires we can bound this term below by something quadratic in $||w-u||$ but this is not possible, so $|w|$ is not strongly convex.
-      \item We calculate Jensen's inequality for $g$: \begin{align*}
+        $\lambda$ is fixed but $w$ was chosen freely so by choosing $w$ arbitrarily close to zero we can obtain a contradiction. To this end, choose $w = \sqrt{\lambda/7}$. Then we have:
+        $$\lambda \leq 3.5\sqrt{\frac{\lambda}{7}}^2 \implies \lambda \leq \frac{\lambda}{2} $$
+        This concludes the proof.
+      \item Let $g = f(w) + \frac{\lambda}{2}||w||^2$, then we can calculate the following: \begin{align*}
         &\hphantom{=} g(\alpha w + (1-\alpha)u, \lambda) + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
         &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}||\alpha w + (1-\alpha) u||^2 + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
-        &\leq f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}(\alpha||w||^2 + (1-\alpha)||u||^2 - \alpha(1-\alpha)||w-u||^2) + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
+        &\leq f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}(\alpha||w||^2 + (1-\alpha)||u||^2 - \alpha(1-\alpha)||w-u||^2) + \\
+        &\hphantom{\leq} \quad \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
         &\leq f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}(\alpha||w||^2 + (1-\alpha)||u||^2) \\
-        &\leq \alpha (f(w) + \frac{\lambda}{2}||w||^2) + (1-\alpha)(f(u) + \frac{\lambda}{2}||u||^2)
+        &\leq \alpha (f(w) + \frac{\lambda}{2}||w||^2) + (1-\alpha)(f(u) + \frac{\lambda}{2}||u||^2) \\
+        &= \alpha g(w) + (1-\alpha)g(u)
       \end{align*}
-        as required.
-        \newpage
+        We conclude that $g$ is $\lambda$-strongly convex.
       \item Suppose $f(w) = g(w) - \frac{\lambda}{2}||w||^2$ is convex. We will show that $g$ is $\lambda$-strongly convex:
@@
           &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}(\alpha w + (1-\alpha) u)^T(\alpha w + (1-\alpha) u) + \\
-          &\hphantom{=} \frac{\lambda}{2}\alpha(1-\alpha)(w-u)^T(w-u) \\
+          &\hphantom{=} \quad \frac{\lambda}{2}\alpha(1-\alpha)(w-u)^T(w-u) \\
           &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}\left(\alpha^2||w||^2+2\alpha(1-\alpha)(u\cdot w) + (1-\alpha)^2||u||^2\right) + \\
-          &\hphantom{=} \frac{\lambda}{2}\alpha(1-\alpha)\left(||w||^2 - 2(u\cdot w) + ||u||^2\right) \\
+          &\hphantom{=} \quad\frac{\lambda}{2}\alpha(1-\alpha)\left(||w||^2 - 2(u\cdot w) + ||u||^2\right) \\
           &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}\left(\alpha^2||w||^2 + (1-\alpha)^2||u||^2\right) + \frac{\lambda}{2}\alpha(1-\alpha)\left(||w||^2 + ||u||^2\right) \\
@@
 
-    \item Exercise 3.7 (Algorithmic Stability for Ridge Regression updated for clarity). Write,
+\begin{problem}{Exercise 3.7 (Algorithmic Stability for Ridge Regression)}
+Write,
     \[
-    f(w, S) = L_S(w) = \frac{1}{m} \sum_{i=1}^{m} (w \cdot x_i - y_i)^2
+    f(w, S) = \hat{L}_S(w) = \frac{1}{m} \sum_{i=1}^{m} (w \cdot x_i - y_i)^2
     \]
@@
         \item Explain why $g(w, S)$ is $\lambda$-strongly convex (as a function of $w$).
-
-        \[\] % Space to write solution
 
@@
 
-\bibliography{sample.bib} % The file containing the bibliography
+%% \bibliography{sample.bib} % The file containing the bibliography
 

### 2023-06-25 assignment-3/article.tex
Solve last 2 problems

@@
       where the last inequality comes from the fact that $f(w_i,S_i) = \frac{1}{m}\ell(w_i, z_i)$.
-      \item We begin by calculating an upper bound for $\ell(w,z)$: \begin{align*}
-        \ell(w, z) &= (w \cdot x - y)^2 \\
-        &= ||w\cdot x||^2 - 2 (w \cdot x) \cdot y + ||y||^2 \\
-        &\leq ||w||^2 ||x||^2 - 2||w|| ||x|| ||y|| + ||y||^2 \\
-        &\leq ||w||^2 ||x||^2 + ||y||^2 \\
-        &\leq C_w^2C_x^2 + C_y^2
-        \end{align*}
-
-        Now we use the hint:
+    \item We begin by applying the hint with $\ell(w_2,z_1) - \ell(w_1,z_1)$:
         \begin{align*}
@@
           \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &= (w_2\cdot x - y - w_1\cdot x + y)(w_2\cdot x - y + w_1\cdot x - y) \\
-          \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &= ((w_2 - w_1)\cdot x)((w_1 + w_2)\cdot x - 2y)
+          \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &= ((w_2 - w_1)\cdot x)((w_1 + w_2)\cdot x - 2y) \\
+          \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &\leq (2C_wC_x)(2C_wC_x+2C_y) \\
+          \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &\leq 2C_wC_x(C_wC_x + C_y)
         \end{align*}
+        Now we can apply the triangle inequality to the right hand side of the inequality from the previous exercise, and then we will apply our upper bound.
+        \begin{align*}
+          &\hphantom{\leq} \frac{1}{m} |\ell(w_2,z_1)-\ell(w_1,z_1)+\ell(w_1,z_2)-\ell(w_2,z_2)| \\
+          &\leq \frac{1}{m} |\ell(w_2,z_1)-\ell(w_1,z_1)| + \frac{1}{m}|\ell(w_1,z_2)-\ell(w_2,z_2)| \\
+          &\leq \frac{4C_wC_x(C_wC_x + C_y)}{m}
+        \end{align*}
+      \item Since $L(A(Z_1)) - L(A(Z_2))$ can be bounded by function of $m$, the Ridge Regression problem is replace-one-stable (ROS).
+
+        The ROS rate, denoted $\beta(m)$ is given by the following:
+        \[
+        \begin{aligned}[t]
+        L(A(Z_1))-L(A(Z_2)) &= \frac{1}{m} |\ell(w_2,z_1)-\ell(w_1,z_1)+\ell(w_1,z_2)-\ell(w_2,z_2)| \\
+        &\leq \frac{4C_wC_x(C_wC_x + C_y)}{m} = \beta(m)
+        \end{aligned}
+        \]
   \end{enumerate}

### 2023-06-25 assignment-3/article.tex
Apply linter changes

@@
 \subsection*{Exercise 3.1 (Rademacher identities)}
-\begin{problem}{Mohri Exercise 3.8. part (c) only}
+\begin{problem}{Mohri Exercise 3.8.\ part (c) only}
 Fix $m \geq 1$. Prove the following identity for any $\alpha \in \mathbb{R}$ and any two hypothesis sets $\mathcal{H}$ and $\mathcal{H}_0$ of functions mapping from $\mathcal{X}$ to $\mathbb{R}$:
 
-$\mathfrak{R}_m(\{\max(h, h_0 ): h \in \mathcal{H}, h_0 \in \mathcal{H}_0 \}) \leq \mathfrak{R}_m (\mathcal{H}) + \mathfrak{R}_m (\mathcal{H}_0)$, where $\max(h, h_0 )$ denotes the function $x \mapsto max_{x\in \mathcal{X}} (h(x), h_0 (x))$ (Hint: you could use the identity $\max(a, b) = \frac{1}{2} [a + b + |a - b|]$ valid for all $a, b \in \mathbb{R}$ and Talagrand’s contraction lemma (see lemma 5.7)).
+$\mathfrak{R}_m(\{\max(h, h_0 ): h \in \mathcal{H}, h_0 \in \mathcal{H}_0 \}) \leq \mathfrak{R}_m (\mathcal{H}) + \mathfrak{R}_m (\mathcal{H}_0)$, where $\max(h, h_0 )$ denotes the function $x \mapsto \max_{x\in \mathcal{X}} (h(x), h_0 (x))$ (Hint: you could use the identity $\max(a, b) = \frac{1}{2} [a + b + |a - b|]$ valid for all $a, b \in \mathbb{R}$ and Talagrand’s contraction lemma (see lemma 5.7)).
 \end{problem}
@@
 \begin{problem}{Mohri Exercise 3.17}
-Show that the VC-dimension of the set of all closed balls in $\mathbb{R}_n$, i.e.
-sets of the form ${x \in \mathbb{R}_n : ||x-x_0|| \leq r}$ for some $x_0 \in \mathbb{R}_n$ and $r \geq 0$, is less than or equal to $n + 2$.
+Show that the VC-dimension of the set of all closed balls in $\mathbb{R}_n$, i.e.\ sets of the form ${x \in \mathbb{R}_n : ||x-x_0|| \leq r}$ for some $x_0 \in \mathbb{R}_n$ and $r \geq 0$, is less than or equal to $n + 2$.
 \end{problem}
@@
   ||x-a||^2 - r &< 0 \\
-  (x-a)^T(x-a) - r &< 0 \\
+  {(x-a)}^T(x-a) - r &< 0 \\
   ||x||^2 - 2(x\cdot a) + ||a||^2 - r &< 0 \\
@@
     -2a_1 \\
-    ... \\
+    \ldots \\
     -2a_n
@@
      x_1 \\
-     ... \\
+     \ldots \\
      x_n
@@
 \begin{problem}{Mohri Exercise 3.20}
-Consider the hypothesis family of sine functions: $\mathcal{H} = \{x \to sin(\omega x): \omega \in \mathbb{R}\}$.
-\begin{enumerate}[label=(\alph*)]
+Consider the hypothesis family of sine functions: $\mathcal{H} = \{x \to \sin(\omega x): \omega \in \mathbb{R}\}$.
+\begin{enumerate}[label= (\alph*)]
 \item Show that for any $x \in \mathbb{R}$ the points $x$, $2x$, $3x$ and $4x$ cannot be shattered by this family of sine functions.
@@
 \begin{solution}
-\begin{enumerate}[label=(\alph*)]
+\begin{enumerate}[label= (\alph*)]
   \item Fix $x\in\mathbb{R}$.
@@
 
-\item Let $m \in \mathbb{N}$ and consider the points $\{2^{-i}: i \leq m\}$ with arbitrary labels $(y_1, ...
-y_{m}) \in \{-1, +1\}^m$.
+\item Let $m \in \mathbb{N}$ and consider the points $\{2^{-i}: i \leq m\}$ with arbitrary labels $(y_1, \ldots y_{m}) \in {\{-1, +1\}}^m$.
 Let $y_i^\prime = \frac{1-y_i}{2}$ and choose $\omega = \pi(1 + \sum_{i=1}^m 2^{i}y_i^\prime)$.
 We will show that this is sufficient to correctly classify our chosen labels.
-then for any $j\in[1,m]$ we have the following:
+Then for any $j\in[1,m]$ we have the following:
     \begin{align*}
@@
       &= \pi \left[2^{-j} + y_j^\prime + \sum_{i=1}^{j-1} 2^{i-j}y_i^\prime + \sum_{i=j+1}^m 2^{i-j}y_i^\prime \right] \\
-      &= \pi \left[2^{-j} + y_j^\prime + \sum_{i=1}^{j-1} 2^{-i}y_{j-i}^\prime + \sum_{i=1}^{m-j} 2^iy_{i+j}^\prime \right] \\
+      &= \pi \left[2^{-j} + y_j^\prime + \sum_{i=1}^{j-1} 2^{-i}y_{j-i}^\prime + \sum_{i=1}^{m-j} 2^i y_{i+j}^\prime \right] \\
       &= \pi \left[2^{-j} + y_j^\prime + \sum_{i=1}^{j-1} 2^{-i}y_{j-i}^\prime\right] + 2\pi \sum_{i=1}^{m-j} 2^{i-1}y_{i+j}^\prime
@@
 Recall that a function $f$ is $\lambda$-strongly convex if for all $u, w \in \mathbb{R}^d , \alpha \in [0, 1]$:
-$$f(\alpha w + (1 - \alpha)u) + \frac{\lambda}{2} \alpha(1 - \alpha)\|w - u\|^2 \leq \alpha f(w) + (1 - \alpha)f(u)$$
+\[f(\alpha w + (1 - \alpha)u) + \frac{\lambda}{2} \alpha(1 - \alpha)\|w - u\|^2 \leq \alpha f(w) + (1 - \alpha)f(u)\]
 \begin{problem}{Exercise 3.5 (Strong convexity)}
-  \begin{enumerate}[label=(\alph*)]
+  \begin{enumerate}[label= (\alph*)]
     \item Prove that $f(w) = \|w\|^2$ is $\lambda$-strongly convex with $\lambda = 2$.
@@
 \begin{solution}
-  \begin{enumerate}[label=(\alph*)]
+  \begin{enumerate}[label= (\alph*)]
       \item We begin by calculating the left hand side of the strong convexity inequality:
         \begin{align*}
-        &\hphantom{=} ||\alpha w + (1-\alpha)u)||^2 + \alpha(1-\alpha)||w-u||^2 \\
-        &= (\alpha w + (1-\alpha)u)^T(\alpha w + (1-\alpha)u) + \alpha(1-\alpha)(w-u)^T(w-u) \\
-        &= \alpha^2||w||^2 + 2\alpha(1-\alpha)(u\cdot w) + (1-\alpha)^2||u||^2 + \alpha(1-\alpha)(||w||^2 - 2(u\cdot w) + ||u||^2) \\
-        &= (\alpha^2+\alpha(1-\alpha))||w||^2 + ((1-\alpha)^2 + \alpha(1-\alpha))||u||^2 \\
+        &\hphantom{=} ||\alpha w + (1-\alpha)u||^2 + \alpha(1-\alpha)||w-u||^2 \\
+        &= {(\alpha w + (1-\alpha)u)}^T(\alpha w + (1-\alpha)u) + \alpha(1-\alpha){(w-u)}^T(w-u) \\
+        &= \alpha^2||w||^2 + 2\alpha(1-\alpha)(u\cdot w) + {(1-\alpha)}^2||u||^2 + \alpha(1-\alpha)(||w||^2 - 2(u\cdot w) + ||u||^2) \\
+        &= (\alpha^2+\alpha(1-\alpha))||w||^2 + ({(1-\alpha)}^2 + \alpha(1-\alpha))||u||^2 \\
         &= \alpha||w||^2 + (1-\alpha)||u||^2
@@
         \begin{align*}
-          \left(\frac{w}{2}\right)^4 + \frac{\lambda}{8}w^2 &\leq \frac{1}{2}w^4 \\
+          {\left(\frac{w}{2}\right)}^4 + \frac{\lambda}{8}w^2 &\leq \frac{1}{2}w^4 \\
           \lambda &\leq \frac{8}{w^2}\left(\frac{w^4}{2} - \frac{w^4}{16}\right) = 3.5w^2
@@
         $\lambda$ is fixed but $w$ was chosen freely so by choosing $w$ arbitrarily close to zero we can obtain a contradiction. To this end, choose $w = \sqrt{\lambda/7}$. Then we have:
-        $$\lambda \leq 3.5\sqrt{\frac{\lambda}{7}}^2 \implies \lambda \leq \frac{\lambda}{2} $$
+        \[\lambda \leq 3.5\sqrt{\frac{\lambda}{7}}^2 \implies \lambda \leq \frac{\lambda}{2}\]
         This concludes the proof.
@@
           &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}||\alpha w + (1-\alpha) u||^2 + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
-          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}(\alpha w + (1-\alpha) u)^T(\alpha w + (1-\alpha) u) + \\
-          &\hphantom{=} \quad \frac{\lambda}{2}\alpha(1-\alpha)(w-u)^T(w-u) \\
-          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}\left(\alpha^2||w||^2+2\alpha(1-\alpha)(u\cdot w) + (1-\alpha)^2||u||^2\right) + \\
+          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}{(\alpha w + (1-\alpha) u)}^T(\alpha w + (1-\alpha) u) + \\
+          &\hphantom{=} \quad \frac{\lambda}{2}\alpha(1-\alpha){(w-u)}^T(w-u) \\
+          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}\left(\alpha^2||w||^2+2\alpha(1-\alpha)(u\cdot w) + {(1-\alpha)}^2||u||^2\right) + \\
           &\hphantom{=} \quad\frac{\lambda}{2}\alpha(1-\alpha)\left(||w||^2 - 2(u\cdot w) + ||u||^2\right) \\
-          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}\left(\alpha^2||w||^2 + (1-\alpha)^2||u||^2\right) + \frac{\lambda}{2}\alpha(1-\alpha)\left(||w||^2 + ||u||^2\right) \\
-          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}||w||^2\left(\alpha^2 + \alpha(1-\alpha)\right) + \frac{\lambda}{2}||u||^2\left((1-\alpha)^2 + \alpha(1-\alpha)\right) \\
+          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}\left(\alpha^2||w||^2 + {(1-\alpha)}^2||u||^2\right) + \frac{\lambda}{2}\alpha(1-\alpha)\left(||w||^2 + ||u||^2\right) \\
+          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}||w||^2\left(\alpha^2 + \alpha(1-\alpha)\right) + \frac{\lambda}{2}||u||^2\left({(1-\alpha)}^2 + \alpha(1-\alpha)\right) \\
           &= f(\alpha w + (1-\alpha)u) + \frac{\lambda\alpha}{2}||w||^2 + \frac{\lambda(1-\alpha)}{2}||u||^2 \\
@@
           &\leq \alpha g(w) + (1-\alpha)g(u) - \frac{\lambda}{2}||\alpha w + (1-\alpha) u||^2 - \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
-          &\leq \alpha g(w) + (1-\alpha)g(u) - \frac{\lambda}{2}(\alpha w + (1-\alpha) u)^T(\alpha w + (1-\alpha) u) - \\
-          &\hphantom{\leq} \frac{\lambda}{2}\alpha(1-\alpha)(w-u)^T(w-u) \\
+          &\leq \alpha g(w) + (1-\alpha)g(u) - \frac{\lambda}{2}{(\alpha w + (1-\alpha) u)}^T(\alpha w + (1-\alpha) u) - \\
+          &\hphantom{\leq} \frac{\lambda}{2}\alpha(1-\alpha){(w-u)}^T(w-u) \\
           &\leq \alpha g(w) + (1-\alpha)g(u) - \frac{\lambda}{2}\left(\alpha^2 ||w||^2 + 2\alpha(1-\alpha)(u \cdot w) + (1-\alpha) ||u||^2\right) - \\
@@
           &\leq \alpha g(w) + (1-\alpha)g(u) - \frac{\lambda}{2}\left(\alpha^2 ||w||^2 + (1-\alpha) ||u||^2\right) - \frac{\lambda}{2}\alpha(1-\alpha)(||w||^2 + ||u||^2) \\
-          &\leq \alpha g(w) - \frac{\lambda}{2}||w||^2\left(\alpha^2 + \alpha(1-\alpha) \right)  + (1-\alpha)g(u) - \frac{\lambda}{2}||u||^2\left((1-\alpha)^2 + \alpha(1-\alpha)\right) \\
+          &\leq \alpha g(w) - \frac{\lambda}{2}||w||^2\left(\alpha^2 + \alpha(1-\alpha) \right)  + (1-\alpha)g(u) - \frac{\lambda}{2}||u||^2\left({(1-\alpha)}^2 + \alpha(1-\alpha)\right) \\
           &\leq \alpha g(w) - \frac{\lambda\alpha}{2}||w||^2  + (1-\alpha)g(u) - \frac{\lambda(1-\alpha)}{2}||u||^2 \\
@@
   \begin{enumerate}
-      \item Find $E[X]^2$, $E[X^2]$, $Var(X)$. Show that the expression $E[X^2] - (E[X])^2 = Var(X)$ corresponds to
+      \item Find ${E[X]}^2$, $E[X^2]$, $Var(X)$. Show that the expression $E[X^2] - {E[X]}^2 = Var(X)$ corresponds to
       \[
@@
       \begin{enumerate}
-        \item $E[X] = \alpha w + (1-\alpha)u$ therefore $E[X]^2 = ||\alpha w + (1-\alpha)u||^2$
+        \item $E[X] = \alpha w + (1-\alpha)u$ therefore ${E[X]}^2 = ||\alpha w + (1-\alpha)u||^2$
         \item $E[X^2] = \alpha||w||^2 + (1-\alpha) ||u||^2$
         \item $\begin{aligned}[t]
-          Var[X] &= E[X^2] - E[X]^2 \\
+          Var[X] &= E[X^2] - {E[X]}^2 \\
           &= \alpha||w||^2 + (1-\alpha) ||u||^2 - ||\alpha w + (1-\alpha)u||^2 \\
-          &= \alpha||w||^2 + (1-\alpha) ||u||^2 - (\alpha w + (1-\alpha)u)^T(\alpha w + (1-\alpha)u) \\
-          &= \alpha||w||^2 + (1-\alpha) ||u||^2 - \left(\alpha^2||w||^2 -2\alpha(1-\alpha)(u\cdot w) + (1-\alpha)^2||u||^2\right) \\
-          &= ||w||^2(\alpha -\alpha^2)+2\alpha(1-\alpha)(u\cdot w) + ||u||^2(1-\alpha-(1-\alpha)^2) \\
+          &= \alpha||w||^2 + (1-\alpha) ||u||^2 - {(\alpha w + (1-\alpha)u)}^T(\alpha w + (1-\alpha)u) \\
+          &= \alpha||w||^2 + (1-\alpha) ||u||^2 - \left(\alpha^2||w||^2 -2\alpha(1-\alpha)(u\cdot w) + {(1-\alpha)}^2||u||^2\right) \\
+          &= ||w||^2(\alpha -\alpha^2)+2\alpha(1-\alpha)(u\cdot w) + ||u||^2(1-\alpha-{(1-\alpha)}^2) \\
           &= ||w||^2\alpha(1 -\alpha)+2\alpha(1-\alpha)(u\cdot w) + ||u||^2\alpha(1 - \alpha) \\
@@
     \[
-    f(w, S) = \hat{L}_S(w) = \frac{1}{m} \sum_{i=1}^{m} (w \cdot x_i - y_i)^2
+    f(w, S) = \hat{L}_S(w) = \frac{1}{m} \sum_{i=1}^{m} {(w \cdot x_i - y_i)}^2
     \]
@@
         \]
-        for $\ell(w, z) = (w \cdot x - y)^2$.
+        for $\ell(w, z) = {(w \cdot x - y)}^2$.
 
@@
     \item $g$ is $\lambda$-strongly convex by the proof in exercise 3.5(c)
-    \item Since $g$ is $\lambda$-strongly convex, we apply the inequality setting $\alpha=0.5$): $$g(\frac{1}{2} w_1 + \frac{1}{2}w_2, S_1) + \frac{\lambda}{4}||w_1-w_2||^2 \leq \frac{1}{2} g(w_1, S_1) + \frac{1}{2}g(w_2, S_1)$$
+    \item Since $g$ is $\lambda$-strongly convex, we apply the inequality setting $\alpha=0.5$: \[g(\frac{1}{2} w_1 + \frac{1}{2}w_2, S_1) + \frac{\lambda}{4}||w_1-w_2||^2 \leq \frac{1}{2} g(w_1, S_1) + \frac{1}{2}g(w_2, S_1)\]
 
       Because $w_1$ minimises $g$ at $S_1$, we have that $g(w_1, S_1) \leq g(\frac{1}{2} w_1 + \frac{1}{2} w_2, S_1)$.
-So we can substitute $g(w_1, S_1)$ in the left hand side in the following way: $$g(w_1, S_1) + \frac{\lambda}{4}||w_1-w_2||^2 \leq \frac{1}{2}g(w_1, s_1) + \frac{1}{2}g(w_2, S_1)$$
+So we can substitute $g(w_1, S_1)$ in the left hand side in the following way: \[g(w_1, S_1) + \frac{\lambda}{4}||w_1-w_2||^2 \leq \frac{1}{2}g(w_1, s_1) + \frac{1}{2}g(w_2, S_1)\]
 
       Subtracting $g(w_1, S_1)$ from both sides of the inequality, and then multiplying both sides by 2 gives the first desired inequality:
-      $$\frac{\lambda}{2}||w_1-w_2||^2 \leq g(w_2,S_1)-g(w_1, S_1)$$
+      \[\frac{\lambda}{2}||w_1-w_2||^2 \leq g(w_2,S_1)-g(w_1, S_1)\]
       We can obtain the other inequality by the same argument, or via symmetry.
-      $$\frac{\lambda}{2}||w_1-w_2||^2 \leq g(w_1,S_2)-g(w_2, S_2)$$
+      \[\frac{\lambda}{2}||w_1-w_2||^2 \leq g(w_1,S_2)-g(w_2, S_2)\]
     \item We start by rewriting the inequalities from the preceding exercise.
@@
         \lambda ||w_1-w_2||^2 &\leq (g(w_2,z_1)-g(w_1,z_1)+g(w_1,z_2)-g(w_2,z_2)) \\
-        \hphantom{\lambda ||w_1-w_2||^2} &\leq (f(w_2,z_1) + ||w_2||^2 -f(w_1,z_1) - ||w_2||^2 \\
+        \hphantom{\lambda ||w_1-w_2||^2} &\leq f(w_2,z_1) + ||w_2||^2 -f(w_1,z_1) - ||w_2||^2 \\
         \hphantom{\lambda ||w_1-w_2||^2}&+f(w_1,z_2) + ||w_1||^2-f(w_2,z_2) - ||w_1||^2 \\
@@
         \begin{align*}
-          \ell(w_2,z_1) - \ell(w_1,z_1) &= (w_2\cdot x - y)^2 - (w_1\cdot x - y)^2 \\
+          \ell(w_2,z_1) - \ell(w_1,z_1) &= {(w_2\cdot x - y)}^2 - {(w_1\cdot x - y)}^2 \\
           \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &= (w_2\cdot x - y - w_1\cdot x + y)(w_2\cdot x - y + w_1\cdot x - y) \\
           \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &= ((w_2 - w_1)\cdot x)((w_1 + w_2)\cdot x - 2y) \\
-          \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &\leq (2C_wC_x)(2C_wC_x+2C_y) \\
-          \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &\leq 2C_wC_x(C_wC_x + C_y)
+          \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &\leq (2C_w C_x)(2C_w C_x+2C_y) \\
+          \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &\leq 2C_w C_x(C_w C_x + C_y)
         \end{align*}
@@
           &\leq \frac{1}{m} |\ell(w_2,z_1)-\ell(w_1,z_1)| + \frac{1}{m}|\ell(w_1,z_2)-\ell(w_2,z_2)| \\
-          &\leq \frac{4C_wC_x(C_wC_x + C_y)}{m}
+          &\leq \frac{4C_w C_x(C_w C_x + C_y)}{m}
         \end{align*}
@@
         L(A(Z_1))-L(A(Z_2)) &= \frac{1}{m} |\ell(w_2,z_1)-\ell(w_1,z_1)+\ell(w_1,z_2)-\ell(w_2,z_2)| \\
-        &\leq \frac{4C_wC_x(C_wC_x + C_y)}{m} = \beta(m)
+        &\leq \frac{4C_w C_x(C_w C_x + C_y)}{m} = \beta(m)
         \end{aligned}

### 2023-06-25 assignment-4/article.tex
Format assignment 4 with latexindent

@@
 \begin{align*}
-  \epsilon_j &= \frac{x_{t,i,j} - \sqrt{\alpha_t}x_i}{\sqrt{1-\alpha_t}} \\
-   &= \frac{x_{t,i,j} - \sqrt{\alpha_t} \cdot \argmin_{x_k \in S^m} ||x_{t,i,j}-x_k||^2}{\sqrt{1-\alpha_t}} \\
-  &= g^*_t(x_{t,i,j})
+	\epsilon_j & = \frac{x_{t,i,j} - \sqrt{\alpha_t}x_i}{\sqrt{1-\alpha_t}}                                              \\
+	           & = \frac{x_{t,i,j} - \sqrt{\alpha_t} \cdot \argmin_{x_k \in S^m} ||x_{t,i,j}-x_k||^2}{\sqrt{1-\alpha_t}} \\
+	           & = g^*_t(x_{t,i,j})
 \end{align*}
@@
 \begin{align*}
-  \hat{L}_{S_t}(g^*_t) &= \frac{1}{m} \sum_{i=1}^{m} \sum_{j \in J} ||g^*_t (x_{t,i,j} ) - \epsilon_j ||^2 \\
-  &= \frac{1}{m} \sum_{i=1}^{m} \sum_{j \in J} ||\epsilon_j - \epsilon_j ||^2 \\
-  &= \frac{1}{m} \sum_{i=1}^{m} \sum_{j \in J} 0 = 0
+	\hat{L}_{S_t}(g^*_t) & = \frac{1}{m} \sum_{i=1}^{m} \sum_{j \in J} ||g^*_t (x_{t,i,j} ) - \epsilon_j ||^2 \\
+	                     & = \frac{1}{m} \sum_{i=1}^{m} \sum_{j \in J} ||\epsilon_j - \epsilon_j ||^2         \\
+	                     & = \frac{1}{m} \sum_{i=1}^{m} \sum_{j \in J} 0 = 0
 \end{align*}
@@
 \begin{align*}
-  p_X(x) &= p_Z(z) \left|\frac{dz}{dx}\right| \\
-  &= 1 \cdot \left| \frac{1}{2 \sqrt{x}} \right| = \frac{1}{2\sqrt{x}}
+	p_X(x) & = p_Z(z) \left|\frac{dz}{dx}\right|                                 \\
+	       & = 1 \cdot \left| \frac{1}{2 \sqrt{x}} \right| = \frac{1}{2\sqrt{x}}
 \end{align*}
@@
 \begin{align*}
-  p_X(x) &= p_Z(z) \left|\frac{dz}{dx}\right| \\
-  &= \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{1}{2}(\log(x(1-x)))^2 \right] \cdot \frac{|1-2x|}{x(1-x)}
+	p_X(x) & = p_Z(z) \left|\frac{dz}{dx}\right|                                                                 \\
+	       & = \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{1}{2}(\log(x(1-x)))^2 \right] \cdot \frac{|1-2x|}{x(1-x)}
 \end{align*}
@@
 \begin{align*}
-  \phi &= \argmin \left[ \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] \log \frac{\sum_{i=1}^I \delta[x-f[z_i,\phi]]}{\kappa q(x)} dx \right] \\
-  &= \argmin \left[ \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] \log \frac{\sum_{i=1}^I \delta[x-f[z_i,\phi]]}{q(x)} dx - \log \kappa \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] dx \right] \\
-  &= \argmin \left[ \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] \log \frac{\sum_{i=1}^I \delta[x-f[z_i,\phi]]}{q(x)} dx - \log \kappa \right]
+	\phi & = \argmin \left[ \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] \log \frac{\sum_{i=1}^I \delta[x-f[z_i,\phi]]}{\kappa q(x)} dx \right]                                                   \\
+	     & = \argmin \left[ \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] \log \frac{\sum_{i=1}^I \delta[x-f[z_i,\phi]]}{q(x)} dx - \log \kappa \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] dx \right] \\
+	     & = \argmin \left[ \int \sum_{i=1}^I \delta[x-f[z_i,\phi]] \log \frac{\sum_{i=1}^I \delta[x-f[z_i,\phi]]}{q(x)} dx - \log \kappa \right]
 \end{align*}
@@
 \begin{align*}
-  Var[x_t] &= Var[\sqrt{1-\beta_t}\cdot x_{t-1} + \sqrt{\beta_t} \cdot \epsilon_t] \\
-  &= Var[\sqrt{1-\beta_t}\cdot x_{t-1}] + Var[\sqrt{\beta_t} \cdot \epsilon_t] \\
-  &= (1-\beta_t) Var[x_{t-1}] + \beta_t Var[ \cdot \epsilon_t] \\
-  &= (1-\beta_t) \textbf{I} + \beta_t \textbf{I} \\
-  &= \textbf{I}
+	Var[x_t] & = Var[\sqrt{1-\beta_t}\cdot x_{t-1} + \sqrt{\beta_t} \cdot \epsilon_t]      \\
+	         & = Var[\sqrt{1-\beta_t}\cdot x_{t-1}] + Var[\sqrt{\beta_t} \cdot \epsilon_t] \\
+	         & = (1-\beta_t) Var[x_{t-1}] + \beta_t Var[ \cdot \epsilon_t]                 \\
+	         & = (1-\beta_t) \textbf{I} + \beta_t \textbf{I}                               \\
+	         & = \textbf{I}
 \end{align*}
@@
 \begin{align*}
-  \mathbb{E}[z] &= 0 \\
-  Var[z] &= a^2 + b^2
+	\mathbb{E}[z] & = 0         \\
+	Var[z]        & = a^2 + b^2
 \end{align*}
@@
 \begin{align*}
-  Var[z] &= Var[a \epsilon_1 + b \epsilon_2] \\
-  &= a^2 Var[\epsilon_1] + b^2 Var[\epsilon_2] + 2ab Cov[\epsilon_1, \epsilon_2] \\
-  &= a^2 Var[\epsilon_1] + b^2 Var[\epsilon_2] \\
-  &= a^2 + b^2
+	Var[z] & = Var[a \epsilon_1 + b \epsilon_2]                                            \\
+	       & = a^2 Var[\epsilon_1] + b^2 Var[\epsilon_2] + 2ab Cov[\epsilon_1, \epsilon_2] \\
+	       & = a^2 Var[\epsilon_1] + b^2 Var[\epsilon_2]                                   \\
+	       & = a^2 + b^2
 \end{align*}

### 2023-06-25 assignment-4/article.tex
Run linter check on assignment 4

@@
 \begin{equation*}
-	L(\phi) = - \sum_j \log\left(1 - \textrm{sig}(f(g(z_j, \theta), \phi)\right) - \sum_i log\left(\textrm{sig}(f(x_i, \phi))\right)
+	L(\phi) = - \sum_j \log\left(1 - \textrm{sig} (f(g(z_j, \theta), \phi))\right) - \sum_i \log\left(\textrm{sig}(f(x_i, \phi))\right)
 \end{equation*}
@@
 \begin{equation*}
-	L(\phi) = - \sum_j \log(1 - D(\hat{x}_j, \phi)) - \sum_i log(D(x_i, \phi))
+	L(\phi) = - \sum_j \log(1 - D(\hat{x}_j, \phi)) - \sum_i \log(D(x_i, \phi))
 \end{equation*}
@@
 The result is a binary cross-entropy loss where the goal is to distinguish $x_i$ from $\hat{x}_i$.
-Real samples $x_i$ decrease the loss by $log(D(x_i, \phi))$ while generated samples increase the loss by $\log(1-D(\hat{x}_j, \phi))$.
+Real samples $x_i$ decrease the loss by $\log(D(x_i, \phi))$ while generated samples increase the loss by $\log(1-D(\hat{x}_j, \phi))$.
 
@@
 \begin{equation*}
-	L(\theta) = - \sum_j \log\left(1 - \textrm{sig}(f(g(z_j, \theta), \phi)\right)
+	L(\theta) = - \sum_j \log\left(1 - \textrm{sig}(f(g(z_j, \theta), \phi))\right)
 \end{equation*}
@@
 If however the generator only produces a subset of the genuine images, then it is sub-optimal.
-This phenomenon is referred to as as "mode collapse", where the generator learns to generate only a limited variety of samples.
+This phenomenon is referred to as `mode collapse', where the generator learns to generate only a limited variety of samples.
 
@@
 
-For each $t \in T$ , let $S_t = \{x_{t,i,j}\}, i = 1, \ldots , m$ for $j \in J$ be the dataset of t-noisy images, $x_{t,i,j} = \sqrt{\alpha_t} x_i + \sqrt{1 - \alpha_t} \epsilon_j$
+For each $t \in T$, let $S_t = \{x_{t,i,j}\}, i = 1, \ldots , m$ for $j \in J$ be the dataset of t-noisy images, $x_{t,i,j} = \sqrt{\alpha_t} x_i + \sqrt{1 - \alpha_t} \epsilon_j$
 
@@
 
-Minimizing this loss amounts to minimizing $||g_t (x_{t,i,j})-\epsilon_j||^2$ for every $i = 1, ..., m$, and every $j \in J$.
+Minimizing this loss amounts to minimizing $||g_t (x_{t,i,j})-\epsilon_j||^2$ for every $i = 1, \ldots, m$, and every $j \in J$.
 
@@
 
-Suppose there are two different minimizers $g^t$ and $h^t$ of the loss. They both achieve the minimum loss, which is zero. So for all $i = 1, ..., m$ and for all $j \in J$,
+Suppose there are two different minimizers $g^t$ and $h^t$ of the loss. They both achieve the minimum loss, which is zero. So for all $i = 1, \ldots, m$ and for all $j \in J$,
 
@@
 
-Consider transforming a uniform base density defined on $z\in[0,1]$ using the function $x = f[z] = z^2$. Find an expression for the transformed distribution $Pr(x)$.
+Consider transforming a uniform base density defined on $z\in[0,1]$ using the function $x = f[z] = z^2$. Find an expression for the transformed distribution $\Pr(x)$.
 
@@
 \begin{equation*}
-	Pr(z) = \frac{1}{\sqrt{2\pi}} \exp \left[\frac{-z^2}{2}\right]
+	\Pr(z) = \frac{1}{\sqrt{2\pi}} \exp \left[\frac{-z^2}{2}\right]
 \end{equation*}
@@
 
-Find an expression for the transformed distribution $Pr(x)$.
+Find an expression for the transformed distribution $\Pr(x)$.
 
@@
 
-As before, let $P_Z = Pr(z)$. Then $x = f[z]$ implies $z = \log(x(1-x))$, from which we derive: $\frac{dz}{dx} = \frac{1-2x}{x(1-x)}$
+As before, let $P_Z = \Pr(z)$. Then $x = f[z]$ implies $z = \log(x(1-x))$, from which we derive: $\frac{dz}{dx} = \frac{1-2x}{x(1-x)}$
 
@@
 	p_X(x) & = p_Z(z) \left|\frac{dz}{dx}\right|                                                                   \\
-	       & = \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{1}{2}(\log(x(1-x)))^2 \right] \cdot \frac{|1-2x|}{x(1-x)}
+	       & = \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{1}{2}{(\log(x(1-x)))}^2 \right] \cdot \frac{|1-2x|}{x(1-x)}
 \end{align*}
@@
 
-Why does it not matter if we can only evaluate the probability q(x) up to a scaling factor $\kappa$? Does the network have to be invertible to minimize this loss function? Explain your reasoning.
+Why does it not matter if we can only evaluate the probability $q(x)$ up to a scaling factor $\kappa$? Does the network have to be invertible to minimize this loss function? Explain your reasoning.
 
@@
 
-$$z = a \cdot \epsilon_1 + b \cdot \epsilon_2$$
+\[z = a \cdot \epsilon_1 + b \cdot \epsilon_2\]
 

### 2023-06-25 assignment-2/article.tex
Add assignment 2

--- 2023-06-25 assignment-3/article.tex
@@
 
-\title{MATH/COMP 562 ASSIGNMENT 3}
+\title{MATH/COMP 562 ASSIGNMENT 2}
 \author{Caleb Moses}
@@
 
+gin{document}
+
 \maketitle
 
-\section*{Rademacher Complexity}
-\subsection*{Exercise 3.1 (Rademacher identities)}
-\begin{problem}{Mohri Exercise 3.8. part (c) only}
-Fix $m \geq 1$. Prove the following identity for any $\alpha \in \mathbb{R}$ and any two hypothesis sets $\mathcal{H}$ and $\mathcal{H}_0$ of functions mapping from $\mathcal{X}$ to $\mathbb{R}$:
+\section{Rademacher Complexity}
 
-$\mathfrak{R}_m(\{\max(h, h_0 ): h \in \mathcal{H}, h_0 \in \mathcal{H}_0 \}) \leq \mathfrak{R}_m (\mathcal{H}) + \mathfrak{R}_m (\mathcal{H}_0)$, where $\max(h, h_0 )$ denotes the function $x \mapsto max_{x\in \mathcal{X}} (h(x), h_0 (x))$ (Hint: you could use the identity $\max(a, b) = \frac{1}{2} [a + b + |a - b|]$ valid for all $a, b \in \mathbb{R}$ and Talagrand’s contraction lemma (see lemma 5.7)).
+\begin{problem}{Exercise 1.1 (Rademacher complexity of linear hypothesis)}
+Prove the following theorem.
+
+\textbf{(Mohri) Theorem 5.10:} Define $B_r = \{x \in \mathbb{R}^d | ||x|| \leq r\}$.
+Let $S = \{x_1 , \ldots, x_m\} \subset X \subset \mathbb{R}^d$ with $X = B_r$.
+Consider the linear hypotheses $h(w, x) = w \cdot x$ and set $H = \{h(x, w) = w \cdot x | x \in X, w \in B_\lambda\}$.
+Prove that the empirical Rademacher complexity is bounded as follows,
+\[
+	\hat{R}_S(H) \leq \frac{r\lambda}{\sqrt{m}}
+\]
+Hint: refer to class notes, Mohri textbook Theorem 5.10.
 \end{problem}
@@
 \begin{solution}
-Let $\mathcal{H}_{\max} = \{\max(h, h_0 ): h \in \mathcal{H}, h_0 \in \mathcal{H_0} \}$ be the class of functions that take the maximum of the predictions of a function $h$ from $\mathcal{H}$ and a function $h_0$ from $\mathcal{H}_0$.
-
-Now we calculate the Rademacher complexity of $\mathcal{H}_{\max}$:
-
-\begin{align}
-\mathfrak{R}_m(\mathcal{H}_{\max}) &= \mathbb{E} \left[ \sup_{h\in\mathcal{H}, h_0\in\mathcal{H}_0} \frac{1}{m} \sum_{i=1}^m \sigma_i \max(h(z_i), h_0(z_i)) \right] \\
-&= \mathbb{E} \left[ \sup_{h\in\mathcal{H}, h_0\in\mathcal{H}_0} \frac{1}{m} \sum_{i=1}^m \frac{\sigma_i}{2} (h(z_i) + h_0(z_i) + |h(z_i) - h_0(z_i)|) \right] \\
-&\leq \frac{1}{2} \mathbb{E}\left[ \sup_{h\in\mathcal{H}}\frac{1}{m} \sum_{i=1}^m \sigma_i h(z_i)\right] + \mathbb{E}\left[\sup_{h_0\in\mathcal{H}_0}\frac{1}{m} \sum_{i=1}^m \sigma_i h_0(z_i)\right] + \\
-&\qquad \frac{1}{2} \mathbb{E}\left[\sup_{h\in\mathcal{H}, h_0\in\mathcal{H}_0} \frac{1}{m} \sum_{i=1}^m \sigma_i |h(z_i) - h_0(z_i)|\right] \\
-&\leq \frac{1}{2} \mathbb{E}\left[ \sup_{h\in\mathcal{H}}\frac{1}{m} \sum_{i=1}^m \sigma_i h(z_i)\right] + \frac{1}{2}\mathbb{E}\left[\sup_{h_0\in\mathcal{H}_0}\frac{1}{m} \sum_{i=1}^m \sigma_i h_0(z_i)\right] + \\
-&\qquad \frac{1}{2} \mathbb{E}\left[\sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i |h(z_i)|\right] + \frac{1}{2} \mathbb{E}\left[\sup_{h_0\in\mathcal{H}_0} \frac{1}{m} \sum_{i=1}^m \sigma_i |h_0(z_i)|\right] \\
-&\leq \frac{\mathfrak{R}_m(\mathcal{H})}{2} + \frac{\mathfrak{R}_m(\mathcal{H}_0)}{2} + \frac{1 \cdot \mathfrak{R}_m(\mathcal{H})}{2} + \frac{1 \cdot \mathfrak{R}_m(\mathcal{H}_0)}{2} \\
-&= \mathfrak{R}_m(\mathcal{H}) + \mathfrak{R}_m(\mathcal{H}_0)
-\end{align}
-
-(1) is the definition of Rademacher complexity for $\mathcal{H}_{\max}$.
-We take advantage of the hint identity in (2) and then in (3) and (4) we apply the linearity of expectation and sub-additivity of the supremum.
-In (6) we apply the triangle inequality to the absolute difference between $h$ and $h_0$, and use sub-additivity of the supremum and linearity of expectation again to split it into two expectations.
-Then in (7) we use the definition of Rademacher complexity to rewrite the first two terms, and the last two require Talagrand's contraction lemma with the fact that the Lipschitz constant of the absolute value function is 1.
-Simplifying gives the sum of Rademacher complexities of $\mathcal{H}$ and $\mathcal{H}_0$ as desired.
-
+	Here is a solution
 \end{solution}
 
-\section{VC dimension}
+\subsection*{Exercise 1.2}
+Given a hypothesis class $H$, of functions $h : X \rightarrow \mathbb{R}$, and a dataset $S = \{x_1 , \ldots, x_m\} \subset X$.
+Define $\Phi(S) = \sup_{h\in H} |E[h] - \hat{E}_S[h]|$ to be the least upper bound for the generalization gap of a function in $H$.
+Prove that $\hat{E}_S [\Phi(S)] \leq 2\hat{R}_m(H)$.
 
-\subsection*{Exercise 3.2 (VC dimension of finite hypothesis sets)}
-\begin{problem}{Mohri Exercise 3.14}
-VC-dimension of finite hypothesis sets. Show that the VC-dimension of a finite hypothesis set $\mathcal{H}$ is at most $\log_2|\mathcal{H}|$.
-\end{problem}
+\subsection*{Exercise 1.3 (Rademacher Identities)}
+Mohri 3.8(a) and 3.8(b).
 
-\begin{solution}
-  Let $\mathcal{H}$ be a finite hypothesis set that shatters an instance space $X$.
-Then each element of $\mathcal{H}$ corresponds to a possible labelling of $X$ and there are $2^{|X|}$ such labellings.
-Since $\mathcal{H}$ shatters $X$ we require $|\mathcal{H}| \geq 2^{|X|}$, since we need at least as many hypotheses as there are labellings for $X$.
-Equivalently, $\log_2{|\mathcal{H}|} \geq |X|$ so the largest set $X$ that $\mathcal{H}$ can shatter has size $\log_2 |\mathcal{H}|$ which gives the desired upper bound on the VC-dimension.
-\end{solution}
+\section{Convex Learning Problems}
+Refer to Ch 12 of Understanding Machine Learning (Shalev-Shwartz).
 
-\subsection*{Exercise 3.3 (VC dimension of closed balls).}
-\begin{problem}{Mohri Exercise 3.17}
-Show that the VC-dimension of the set of all closed balls in $\mathbb{R}_n$, i.e.
-sets of the form ${x \in \mathbb{R}_n : ||x-x_0|| \leq r}$ for some $x_0 \in \mathbb{R}_n$ and $r \geq 0$, is less than or equal to $n + 2$.
-\end{problem}
+\subsection*{Exercise 1.4 (non convexity of 0\-1 loss)}
+Problem 12.1. Hint: Consider samples in a checkerboard pattern.
 
-\begin{solution}
-Fix $r>0$, then the constraint $||x-a||^2<r$ is equivalent to:
-\begin{align*}
-  ||x-a||^2 - r &< 0 \\
-  (x-a)^T(x-a) - r &< 0 \\
-  ||x||^2 - 2(x\cdot a) + ||a||^2 - r &< 0 \\
-  \sum_{i=1}^n x_i^2 - 2 \sum_{i=1}^n a_i x_i + \sum_{i=1}^n a_i^2 - r &< 0
-\end{align*}
-\end{solution}
+\subsection*{Exercise 1.5 (Convexity, Lipschitz, and Smoothness of logistic regression loss)}
+Problem 12.2.
 
-We will show that there are enough free variables to capture $n+1$ arbitrary points. The last line is equivalent to $WX - B < 0$ with:
-
-\begin{align*}
-  W &= \begin{bmatrix}
-    1 \\
-    -2a_1 \\
-    ... \\
-    -2a_n
-   \end{bmatrix} & X = \begin{bmatrix}
-     \sum_{i=1}^n x_i^2 \\
-     x_1 \\
-     ... \\
-     x_n
-   \end{bmatrix} &&  B = \sum_{i=1}^n r - a_i^2
-\end{align*}
-
-There are n+1 linearly independent equations with n+1 unknowns (1 for each $a_i$ from 1 to n, plus $r$).
-So the VC dimension of closed balls in $\mathbb{R}^n$ is at most equal to the VC dimension of hyperplanes in $\mathbb{R}^{n+1}$, which is $n+2$.
-\newpage
-\subsection*{Exercise 3.4 (Function class with infinite VC dimension).}
-\begin{problem}{Mohri Exercise 3.20}
-Consider the hypothesis family of sine functions: $\mathcal{H} = \{x \to sin(\omega x): \omega \in \mathbb{R}\}$.
-\begin{enumerate}[label=(\alph*)]
-\item Show that for any $x \in \mathbb{R}$ the points $x$, $2x$, $3x$ and $4x$ cannot be shattered by this family of sine functions.
-\item Show that the VC-dimension of the family of sine functions is infinite. (Hint: show that $\{2^{-i}: i \leq m\}$ can be shattered for any $m > 0$.)
-\end{enumerate}
-\end{problem}
-
-\begin{solution}
-\begin{enumerate}[label=(\alph*)]
-  \item Fix $x\in\mathbb{R}$.
-Consider the dichotomy $\{(x, -), (2x, -), (3x, +), (4x, -)\}$.
-Assume towards contradiction that $\mathcal{H}$ realises this dichotomy, so there exists $\omega\in\mathbb{R}$ satisfying $\sin(\omega x) < 0$, $\sin(2 \omega x) < 0$, $\sin(3 \omega x) > 0$ and $\sin(4 \omega x) < 0$.
-
-  Using the double angle formulae for sin and cos gives:
-  \begin{align*}
-    \sin(4\omega x) &= 2 \sin(2\omega x)\cos(2\omega x) \\
-    &= 2\sin(2\omega x) (1 - 2\sin^2(\omega x))
-  \end{align*}
-
-  Since $\sin(4\omega x) < 0$ and $\sin(2\omega x) < 0$ by assumption, $1 - 2\sin^2(\omega x) > 0$ must hold otherwise the right hand side is positive while the left hand side is negative.
-Therefore $\sin^2(\omega x) < \frac{1}{2}$.
-
-  Next we use the triple angle formula for $\sin$ to calculate the following:
-  \begin{align*}
-    \sin(3\omega x) &= \sin(\omega x)(3 - 4\sin^2(\omega x))
-  \end{align*}
-
-  Since $\sin(3\omega x) > 0$ and $\sin(\omega x) < 0$, we require $3 - 4\sin^2(\omega x) < 0$ otherwise the left hand side is positive and the right hand side is negative.
-Therefore $\sin^2(\omega x) > \frac{3}{4}$.
-This contradicts $\sin^2(\omega x) < \frac{1}{2}$, so there is no $\omega$ such that $\mathcal{H}$ shatters $x$, $2x$, $3x$ and $4x$.
-
-\item Let $m \in \mathbb{N}$ and consider the points $\{2^{-i}: i \leq m\}$ with arbitrary labels $(y_1, ...
-y_{m}) \in \{-1, +1\}^m$.
-Let $y_i^\prime = \frac{1-y_i}{2}$ and choose $\omega = \pi(1 + \sum_{i=1}^m 2^{i}y_i^\prime)$.
-We will show that this is sufficient to correctly classify our chosen labels.
-then for any $j\in[1,m]$ we have the following:
-    \begin{align*}
-      \omega 2^{-j} &= 2^{-j} \pi\left[1 + \sum_{i=1}^m 2^{i}y_i^\prime\right] \\
-      &= \pi \left[2^{-j} + \sum_{i=1}^m 2^{i-j}y_i^\prime\right] \\
-      &= \pi \left[2^{-j} + y_j^\prime + \sum_{i=1}^{j-1} 2^{i-j}y_i^\prime + \sum_{i=j+1}^m 2^{i-j}y_i^\prime \right] \\
-      &= \pi \left[2^{-j} + y_j^\prime + \sum_{i=1}^{j-1} 2^{-i}y_{j-i}^\prime + \sum_{i=1}^{m-j} 2^iy_{i+j}^\prime \right] \\
-      &= \pi \left[2^{-j} + y_j^\prime + \sum_{i=1}^{j-1} 2^{-i}y_{j-i}^\prime\right] + 2\pi \sum_{i=1}^{m-j} 2^{i-1}y_{i+j}^\prime
-    \end{align*}
-  The last term contributes multiples of $2\pi$ to $\omega 2^{-i}$ and so can be ignored. Since $y_i\in\{0,1\}$, the remaining term can be bound in the following way:
-
-  \begin{align*}
-    \pi \left[2^{-j} + y_j^\prime + \sum_{i=1}^{j-1} 2^{-i}y_{j-i}^\prime\right] &\leq \pi \left[ y_j^\prime + \sum_{i=1}^{j} 2^{-i}\right] < \pi(1 + y_j^\prime) \\
-    \pi \left[2^{-j} + y_j^\prime + \sum_{i=1}^{j-1} 2^{-i}y_{j-i}^\prime\right] &> \pi y_j^\prime
-  \end{align*}
-
-  Now if $y_j = 1$ then $y_j^\prime = 0$ and $0 < \omega 2^{-j} < \pi$ which implies $\sin(\omega 2^{-j}) > 0$.
-Similarly, if $y_j = -1$, then $y_j^\prime = 1$, $\pi < \omega 2^{-j} < 2\pi$ and we have that $\sin(\omega 2^{-j}) < 0$ as desired.
-
-\end{enumerate}
-\end{solution}
-
-\section*{Stability Theory}
-Recall that a function $f$ is $\lambda$-strongly convex if for all $u, w \in \mathbb{R}^d , \alpha \in [0, 1]$:
-$$f(\alpha w + (1 - \alpha)u) + \frac{\lambda}{2} \alpha(1 - \alpha)\|w - u\|^2 \leq \alpha f(w) + (1 - \alpha)f(u)$$
-\begin{problem}{Exercise 3.5 (Strong convexity)}
-  \begin{enumerate}[label=(\alph*)]
-    \item Prove that $f(w) = \|w\|^2$ is $\lambda$-strongly convex with $\lambda = 2$.
-    \item Give an example of a convex function which is not $\lambda$-strongly convex for any $\lambda > 0$.
-    \item Let $f : \mathbb{R}^d \rightarrow \mathbb{R}$ be convex. Prove that $g(w, \lambda) = f(w) + \frac{\lambda}{2}\|w\|^2$ is $\lambda$-strongly convex.
-    \item Prove that $g$ is $\lambda$-strongly convex if and only if $f(w) = g(w) - \frac{\lambda}{2}\|w\|^2$ is convex.
-  \end{enumerate}
-\end{problem}
-
-\begin{solution}
-  \begin{enumerate}[label=(\alph*)]
-      \item We begin by calculating the left hand side of the strong convexity inequality:
-        \begin{align*}
-        &\hphantom{=} ||\alpha w + (1-\alpha)u)||^2 + \alpha(1-\alpha)||w-u||^2 \\
-        &= (\alpha w + (1-\alpha)u)^T(\alpha w + (1-\alpha)u) + \alpha(1-\alpha)(w-u)^T(w-u) \\
-        &= \alpha^2||w||^2 + 2\alpha(1-\alpha)(u\cdot w) + (1-\alpha)^2||u||^2 + \alpha(1-\alpha)(||w||^2 - 2(u\cdot w) + ||u||^2) \\
-        &= (\alpha^2+\alpha(1-\alpha))||w||^2 + ((1-\alpha)^2 + \alpha(1-\alpha))||u||^2 \\
-        &= \alpha||w||^2 + (1-\alpha)||u||^2
-        \end{align*}
-      \item I claim that $f(x) = x^4$ is convex, but not $\lambda$-strongly convex for any $\lambda>0$. We know $f$ is convex since $f^{\prime\prime}(x) = 12x^2 > 0$ for all x.
-        Now to show $f$ is not $\lambda$-strongly convex, choose $x_0=0$, $x_1=w$ and $\alpha=\frac{1}{2}$.
-Suppose towards contradiction that $f$ is $\lambda$-strongly convex for some $\lambda > 0$.
-Then the following must hold for all $w\in\mathbb{R}$:
-        \begin{align*}
-          \left(\frac{w}{2}\right)^4 + \frac{\lambda}{8}w^2 &\leq \frac{1}{2}w^4 \\
-          \lambda &\leq \frac{8}{w^2}\left(\frac{w^4}{2} - \frac{w^4}{16}\right) = 3.5w^2
-        \end{align*}
-        $\lambda$ is fixed but $w$ was chosen freely so by choosing $w$ arbitrarily close to zero we can obtain a contradiction. To this end, choose $w = \sqrt{\lambda/7}$. Then we have:
-        $$\lambda \leq 3.5\sqrt{\frac{\lambda}{7}}^2 \implies \lambda \leq \frac{\lambda}{2} $$
-        This concludes the proof.
-      \item Let $g = f(w) + \frac{\lambda}{2}||w||^2$, then we can calculate the following: \begin{align*}
-        &\hphantom{=} g(\alpha w + (1-\alpha)u, \lambda) + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
-        &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}||\alpha w + (1-\alpha) u||^2 + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
-        &\leq f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}(\alpha||w||^2 + (1-\alpha)||u||^2 - \alpha(1-\alpha)||w-u||^2) + \\
-        &\hphantom{\leq} \quad \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
-        &\leq f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}(\alpha||w||^2 + (1-\alpha)||u||^2) \\
-        &\leq \alpha (f(w) + \frac{\lambda}{2}||w||^2) + (1-\alpha)(f(u) + \frac{\lambda}{2}||u||^2) \\
-        &= \alpha g(w) + (1-\alpha)g(u)
-      \end{align*}
-        We conclude that $g$ is $\lambda$-strongly convex.
-      \item Suppose $f(w) = g(w) - \frac{\lambda}{2}||w||^2$ is convex. We will show that $g$ is $\lambda$-strongly convex:
-        \begin{align*}
-          &\hphantom{=} g(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
-          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}||\alpha w + (1-\alpha) u||^2 + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
-          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}(\alpha w + (1-\alpha) u)^T(\alpha w + (1-\alpha) u) + \\
-          &\hphantom{=} \quad \frac{\lambda}{2}\alpha(1-\alpha)(w-u)^T(w-u) \\
-          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}\left(\alpha^2||w||^2+2\alpha(1-\alpha)(u\cdot w) + (1-\alpha)^2||u||^2\right) + \\
-          &\hphantom{=} \quad\frac{\lambda}{2}\alpha(1-\alpha)\left(||w||^2 - 2(u\cdot w) + ||u||^2\right) \\
-          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}\left(\alpha^2||w||^2 + (1-\alpha)^2||u||^2\right) + \frac{\lambda}{2}\alpha(1-\alpha)\left(||w||^2 + ||u||^2\right) \\
-          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda}{2}||w||^2\left(\alpha^2 + \alpha(1-\alpha)\right) + \frac{\lambda}{2}||u||^2\left((1-\alpha)^2 + \alpha(1-\alpha)\right) \\
-          &= f(\alpha w + (1-\alpha)u) + \frac{\lambda\alpha}{2}||w||^2 + \frac{\lambda(1-\alpha)}{2}||u||^2 \\
-          &\leq \alpha f(w) + (1-\alpha)f(u) + \frac{\lambda\alpha}{2}||w||^2 + \frac{\lambda(1-\alpha)}{2}||u||^2 \\
-          &\leq \alpha (f(w) + \frac{\lambda}{2}||w||^2) + (1-\alpha)(f(u) + \frac{\lambda}{2}||u||^2) \\
-          &= \alpha g(w) + (1-\alpha)g(u)
-        \end{align*}
-        as desired.
-
-        Now we must prove the converse. To that aim, assume $g$ is $\lambda$-strongly convex. We will now prove that $f$ is convex.
-        \begin{align*}
-          &\hphantom{=} f(\alpha w + (1-\alpha)u) \\
-          &= g(\alpha w + (1-\alpha)u) - \frac{\lambda}{2}||\alpha w + (1-\alpha)u||^2 \\
-          &\leq \alpha g(w) + (1-\alpha)g(u) - \frac{\lambda}{2}||\alpha w + (1-\alpha) u||^2 - \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
-          &\leq \alpha g(w) + (1-\alpha)g(u) - \frac{\lambda}{2}(\alpha w + (1-\alpha) u)^T(\alpha w + (1-\alpha) u) - \\
-          &\hphantom{\leq} \frac{\lambda}{2}\alpha(1-\alpha)(w-u)^T(w-u) \\
-          &\leq \alpha g(w) + (1-\alpha)g(u) - \frac{\lambda}{2}\left(\alpha^2 ||w||^2 + 2\alpha(1-\alpha)(u \cdot w) + (1-\alpha) ||u||^2\right) - \\
-          &\hphantom{\leq} \frac{\lambda}{2}\alpha(1-\alpha)(||w||^2 - 2(u\cdot w) + ||u||^2) \\
-          &\leq \alpha g(w) + (1-\alpha)g(u) - \frac{\lambda}{2}\left(\alpha^2 ||w||^2 + (1-\alpha) ||u||^2\right) - \frac{\lambda}{2}\alpha(1-\alpha)(||w||^2 + ||u||^2) \\
-          &\leq \alpha g(w) - \frac{\lambda}{2}||w||^2\left(\alpha^2 + \alpha(1-\alpha) \right)  + (1-\alpha)g(u) - \frac{\lambda}{2}||u||^2\left((1-\alpha)^2 + \alpha(1-\alpha)\right) \\
-          &\leq \alpha g(w) - \frac{\lambda\alpha}{2}||w||^2  + (1-\alpha)g(u) - \frac{\lambda(1-\alpha)}{2}||u||^2 \\
-          &\leq \alpha \left(g(w) - \frac{\lambda}{2}||w||^2\right) + (1-\alpha)\left(g(u) - \frac{\lambda}{2}||u||^2\right) \\
-          &= \alpha f(w) + (1-\alpha)f(u)
-        \end{align*}
-        so $f$ is convex.
-  \end{enumerate}
-\end{solution}
-
-\begin{problem}{Exercise 3.6 (Random Variables and Strong Convexity)}
-  Let $X \in \mathbb{R}$ be a discrete vector-valued random variable, with distribution
-  \[
-  \rho(X) =
-  \begin{cases}
-    w, & \text{with probability }\alpha \\
-    u, & \text{with probability }(1 - \alpha)
-  \end{cases}
-  \]
-  \begin{enumerate}
-      \item Find $E[X]^2$, $E[X^2]$, $Var(X)$. Show that the expression $E[X^2] - (E[X])^2 = Var(X)$ corresponds to
-      \[
-      \alpha f(w) + (1 - \alpha) f(u) - f(\alpha w + (1 - \alpha) u) = \alpha (1 - \alpha) \|w - u\|^2
-      \]
-      with $f(w) = w^2$.
-
-      \item Suppose $f$ is $\lambda$-strongly convex. Let $X$ be as above. Show that
-      \[
-      E[f(X)] - f(E[X]) \geq \frac{\lambda}{2} Var(X)
-      \]
-  \end{enumerate}
-\end{problem}
-
-\begin{solution}
-  \begin{enumerate}
-    \item
-      \begin{enumerate}
-        \item $E[X] = \alpha w + (1-\alpha)u$ therefore $E[X]^2 = ||\alpha w + (1-\alpha)u||^2$
-        \item $E[X^2] = \alpha||w||^2 + (1-\alpha) ||u||^2$
-        \item $\begin{aligned}[t]
-          Var[X] &= E[X^2] - E[X]^2 \\
-          &= \alpha||w||^2 + (1-\alpha) ||u||^2 - ||\alpha w + (1-\alpha)u||^2 \\
-          &= \alpha||w||^2 + (1-\alpha) ||u||^2 - (\alpha w + (1-\alpha)u)^T(\alpha w + (1-\alpha)u) \\
-          &= \alpha||w||^2 + (1-\alpha) ||u||^2 - \left(\alpha^2||w||^2 -2\alpha(1-\alpha)(u\cdot w) + (1-\alpha)^2||u||^2\right) \\
-          &= ||w||^2(\alpha -\alpha^2)+2\alpha(1-\alpha)(u\cdot w) + ||u||^2(1-\alpha-(1-\alpha)^2) \\
-          &= ||w||^2\alpha(1 -\alpha)+2\alpha(1-\alpha)(u\cdot w) + ||u||^2\alpha(1 - \alpha) \\
-          &= \alpha(1-\alpha)\left(||w||^2 - 2(u\cdot w) + ||u||^2\right) \\
-          &= \alpha(1-\alpha)||w-u||^2
-          \end{aligned}$
-      \end{enumerate}
-    \item $\begin{aligned}[t]
-      E[f(X)] - f(E[X]) &= \alpha f(w) + (1-\alpha)f(u) - f(\alpha w + (1-\alpha)u) \\
-      &\geq \alpha f(w) + (1-\alpha)f(u) - \alpha f(w) -(1-\alpha)f(u) + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
-      &\geq \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 = \frac{\lambda}{2}Var[x]
-      \end{aligned}$
-  \end{enumerate}
-\end{solution}
-
-\begin{problem}{Exercise 3.7 (Algorithmic Stability for Ridge Regression)}
-Write,
-    \[
-    f(w, S) = \hat{L}_S(w) = \frac{1}{m} \sum_{i=1}^{m} (w \cdot x_i - y_i)^2
-    \]
-    Consider the datasets $S_1$, $S_2$, which differ by exactly one data point. Write
-    \[
-    z_1 = (x_1, y_1) \in S_1, \quad z_2 = (x_2, y_2) \in S_2
-    \]
-    for the different points. Define $g(w, S) = f(w, S) + \frac{\lambda}{2} \|w\|^2$ and
-    \[
-    w_i = A(S_i) = \arg\min_w g(w, S_i), \quad i = 1, 2
-    \]
-    \begin{enumerate}
-        \item Explain why $g(w, S)$ is $\lambda$-strongly convex (as a function of $w$).
-
-        \item Explain how strong convexity is applied to obtain
-        \[
-        g(w_2, S_1) - g(w_1, S_1) \geq \frac{\lambda}{2} \|w_1 - w_2\|^2
-        \]
-        \[
-        g(w_1, S_2) - g(w_2, S_2) \geq \frac{\lambda}{2} \|w_1 - w_2\|^2
-        \]
-
-        \item Simplify to obtain
-        \[
-        \lambda \|w_1 - w_2\|^2 \leq \frac{1}{m} | \ell(w_2, z_1) - \ell(w_1, z_1) + \ell(w_1, z_2) - \ell(w_2, z_2) |
-        \]
-        for $\ell(w, z) = (w \cdot x - y)^2$.
-
-        \item Assume that $\|x\| \leq C_x$ and $|y| \leq C_y$ for all data $x$, $y$, and that $\|w_1\|, \|w_2\| \leq C_w$.
-Bound the RHS of the previous inequality.
-(Hint: apply the identity $(a^2 - b^2) = (a - b)(a + b)$ to the loss.)
-
-        \item Conclude that the Ridge Regression problem is replace one stable (ROS) and give the rate as a function of $\lambda$, $m$.
-    \end{enumerate}
-\end{problem}
-
-\begin{solution}
-  \begin{enumerate}
-    \item $g$ is $\lambda$-strongly convex by the proof in exercise 3.5(c)
-    \item Since $g$ is $\lambda$-strongly convex, we apply the inequality setting $\alpha=0.5$): $$g(\frac{1}{2} w_1 + \frac{1}{2}w_2, S_1) + \frac{\lambda}{4}||w_1-w_2||^2 \leq \frac{1}{2} g(w_1, S_1) + \frac{1}{2}g(w_2, S_1)$$
-
-      Because $w_1$ minimises $g$ at $S_1$, we have that $g(w_1, S_1) \leq g(\frac{1}{2} w_1 + \frac{1}{2} w_2, S_1)$.
-So we can substitute $g(w_1, S_1)$ in the left hand side in the following way: $$g(w_1, S_1) + \frac{\lambda}{4}||w_1-w_2||^2 \leq \frac{1}{2}g(w_1, s_1) + \frac{1}{2}g(w_2, S_1)$$
-
-      Subtracting $g(w_1, S_1)$ from both sides of the inequality, and then multiplying both sides by 2 gives the first desired inequality:
-      $$\frac{\lambda}{2}||w_1-w_2||^2 \leq g(w_2,S_1)-g(w_1, S_1)$$
-      We can obtain the other inequality by the same argument, or via symmetry.
-      $$\frac{\lambda}{2}||w_1-w_2||^2 \leq g(w_1,S_2)-g(w_2, S_2)$$
-    \item We start by rewriting the inequalities from the preceding exercise.
-      \begin{align*}
-        \frac{\lambda}{2}||w_1-w_2||^2 \leq g(w_2, z_1)-g(w_1, z_1) \\
-        \frac{\lambda}{2}||w_1-w_2||^2 \leq g(w_1, z_2)-g(w_2, z_2)
-      \end{align*}
-      Then we add both inequalities together to obtain:
-      \begin{align*}
-        \lambda ||w_1-w_2||^2 &\leq (g(w_2,z_1)-g(w_1,z_1)+g(w_1,z_2)-g(w_2,z_2)) \\
-        \hphantom{\lambda ||w_1-w_2||^2} &\leq (f(w_2,z_1) + ||w_2||^2 -f(w_1,z_1) - ||w_2||^2 \\
-        \hphantom{\lambda ||w_1-w_2||^2}&+f(w_1,z_2) + ||w_1||^2-f(w_2,z_2) - ||w_1||^2 \\
-        &\leq f(w_2,z_1)-f(w_1,z_1)+f(w_1,z_2)-f(w_2,z_2) \\
-        &\leq \frac{1}{m} (\ell(w_2,z_1)-\ell(w_1,z_1)+\ell(w_1,z_2)-\ell(w_2,z_2))
-      \end{align*}
-      where the last inequality comes from the fact that $f(w_i,S_i) = \frac{1}{m}\ell(w_i, z_i)$.
-      \item We begin by calculating an upper bound for $\ell(w,z)$: \begin{align*}
-        \ell(w, z) &= (w \cdot x - y)^2 \\
-        &= ||w\cdot x||^2 - 2 (w \cdot x) \cdot y + ||y||^2 \\
-        &\leq ||w||^2 ||x||^2 - 2||w|| ||x|| ||y|| + ||y||^2 \\
-        &\leq ||w||^2 ||x||^2 + ||y||^2 \\
-        &\leq C_w^2C_x^2 + C_y^2
-        \end{align*}
-
-        Now we use the hint:
-        \begin{align*}
-          \ell(w_2,z_1) - \ell(w_1,z_1) &= (w_2\cdot x - y)^2 - (w_1\cdot x - y)^2 \\
-          \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &= (w_2\cdot x - y - w_1\cdot x + y)(w_2\cdot x - y + w_1\cdot x - y) \\
-          \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &= ((w_2 - w_1)\cdot x)((w_1 + w_2)\cdot x - 2y)
-        \end{align*}
-  \end{enumerate}
-\end{solution}
+\subsection*{Exercise 1.6 (Lipschitz continuity of the hinge loss)}
+Problem 12.3.
 

### 2023-06-29 assignment-2/article.tex
Tidy up
Whitespace cleanup
Fix problem 1 statement
Solve 1.1

@@
 
-gin{document}
-
 \maketitle
@@
 
-\textbf{(Mohri) Theorem 5.10:} Define $B_r = \{x \in \mathbb{R}^d | ||x|| \leq r\}$.
+\textbf{(Mohri) Theorem 5.10:} Define $B_r = \{x \in \mathbb{R}^d: ||x|| \leq r\}$.
 Let $S = \{x_1 , \ldots, x_m\} \subset X \subset \mathbb{R}^d$ with $X = B_r$.
-Consider the linear hypotheses $h(w, x) = w \cdot x$ and set $H = \{h(x, w) = w \cdot x | x \in X, w \in B_\lambda\}$.
+Consider the linear hypotheses $h(w, x) = w \cdot x$ and set $\mathcal{H} = \{h(x, w) = w \cdot x: x \in X, w \in B_\lambda\}$.
 Prove that the empirical Rademacher complexity is bounded as follows,
 \[
-	\hat{R}_S(H) \leq \frac{r\lambda}{\sqrt{m}}
+  \hat{\mathfrak{R}}_S(H) \leq \frac{r\lambda}{\sqrt{m}}
 \]
@@
 \begin{solution}
-	Here is a solution
+  \begin{align}
+    \hat{\mathfrak{R}}_S(\mathcal{H}) & = \frac{1}{m}\mathbb{E}_\sigma\left[ \sup_{||w||\leq \lambda}\sum_{i=1}^m\sigma_i w \cdot x_i \right] \text{ by definition}                                                                               \\
+                                      & = \frac{1}{m}\mathbb{E}_\sigma\left[ \sup_{||w||\leq \lambda} w\cdot \sum_{i=1}^m \sigma_i x_i \right] \text{ by linearity of the dot product}                                                            \\
+                                      & \leq \frac{1}{m}\mathbb{E}_\sigma\left[ \lambda \left|\left|\sum_{i=1}^m \sigma_i x_i \right|\right|\right] \text{ applying Cauchy-Schwartz and $||w|| \leq \lambda$}                                     \\
+                                      & \leq \frac{\lambda}{m}\mathbb{E}_\sigma\left[ \left|\left|\sum_{i=1}^m \sigma_i x_i \right|\right|\right] \text{ by the linearity of expectation}                                                         \\
+                                      & \leq \frac{\lambda}{m}\mathbb{E}_\sigma\left[ {\left(\left|\left|\sum_{i=1}^m \sigma_i x_i \right|\right|^2\right)}^\frac{1}{2}\right] \text{ since $||x|| = \sqrt{||x||^2}$}                             \\
+                                      & = \frac{\lambda}{m}\mathbb{E}_\sigma\left[ {\left(\sum_{i,j=1}^m \sigma_j \sigma_j (x_i \cdot x_j)\right)}^{\frac{1}{2}} \right] \text{since $||x||^2 = x^T x$ and rearranging terms}                     \\
+                                      & \leq \frac{\lambda}{m}{\left[\sum_{i=1}^m ||x_i||^2\right]}^{\frac{1}{2}} \text{ since $\mathbb{E}_\sigma[\sigma_i \sigma_j] = \mathbb{E}_\sigma[\sigma_i]\mathbb{E}_\sigma[\sigma_j] = 0$ for $i\neq j$} \\
+                                      & \leq \frac{\lambda \sqrt{mr^2}}{m} = \frac{r\lambda}{m} \text{ as required.}
+  \end{align}
+  Note that we are able to remove the expectation in line (7) since the dependency on $\sigma$ is removed.
 \end{solution}

### 2023-06-29 assignment-2/article.tex
Solve 1.2

@@
 
-\subsection*{Exercise 1.2}
-Given a hypothesis class $H$, of functions $h : X \rightarrow \mathbb{R}$, and a dataset $S = \{x_1 , \ldots, x_m\} \subset X$.
-Define $\Phi(S) = \sup_{h\in H} |E[h] - \hat{E}_S[h]|$ to be the least upper bound for the generalization gap of a function in $H$.
-Prove that $\hat{E}_S [\Phi(S)] \leq 2\hat{R}_m(H)$.
+\begin{problem}{Exercise 1.2}
+Given a hypothesis class $\mathcal{H}$, of functions $h : X \rightarrow \mathbb{R}$, and a dataset $S = \{x_1 , \ldots, x_m\} \subset X$.
+Define $\Phi(S) = \sup_{h\in \mathcal{H}} |\mathbb{E}[h] - \hat{\mathbb{E}}_S[h]|$ to be the least upper bound for the generalization gap of a function in $\mathcal{H}$.
 
-\subsection*{Exercise 1.3 (Rademacher Identities)}
-Mohri 3.8(a) and 3.8(b).
+Prove that $\hat{E}_S [\Phi(S)] \leq 2\hat{\mathfrak{R}}_m(\mathcal{H})$.
+\end{problem}
 

### 2023-06-29 assignment-2/article.tex
Solve 1.2
Solve 1.3

@@
 
+\begin{solution}
+  Let $S$ and $S^\prime$ be two datasets differing in one point, say $z_m\in S$ and $z_m^\prime \in S^\prime$. We proceed by bounding the expectation $\hat{\mathbb{E}}_S[\Phi(S)]$ as follows:
+  \begin{align*}
+    \hat{\mathbb{E}}_S[\Phi(S)] & = \hat{\mathbb{E}}_S\left[ \sup_{h\in\mathcal{H}} \left| \mathbb{E}[h] - \hat{\mathbb{E}}_S[h] \right| \right]                                                                                       \\
+                                & = \mathbb{E}_S \left[ \sup_{h\in\mathcal{H}} \hat{\mathbb{E}}_{S^\prime}[h] - \hat{\mathbb{E}}_{S}[h] \right]                                                                                        \\
+                                & \leq \mathbb{E}_{S,S^\prime}\left[ \sup_{h\in\mathcal{H}} \hat{\mathbb{E}}_{S^\prime}[h] - \hat{\mathbb{E}}_{S}[h] \right]                                                                           \\
+                                & = \mathbb{E}_{S,S^\prime} \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m\left(h(z_i^\prime) - h(z_i)\right) \right]                                                                          \\
+                                & = \mathbb{E}_{\sigma, S, S^\prime} \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i (h(z_i^\prime) - h(z_i)) \right]                                                                  \\
+                                & \leq \mathbb{E}_{\sigma, S^\prime}\left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sigma_i h(z_i^\prime) \right] + \mathbb{E}_{\sigma, S}\left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sigma_i h(z_i) \right] \\
+                                & = 2\mathbb{E}_{\sigma,S}\left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma h(z_i) \right] = 2\hat{\mathfrak{R}}_m(\mathcal{H})
+  \end{align*}
+\end{solution}
+
+\begin{problem}{Exercise 1.3 (Rademacher Identities)}
+Fix $m\geq 1$.
+Prove the following identities for any $\alpha \in \mathbb{R}$ and any two hypothesis sets $\mathcal{H}$ and $\mathcal{H^\prime}$ of functions mapping from $\mathcal{X}$ to $\mathcal{R}$.
+\begin{enumerate}[label= (\alph*)]
+  \item $\mathfrak{R}_m(\alpha \mathcal{H}) = |\alpha| \mathfrak{R}_m(\mathcal{H})$.
+  \item $\mathfrak{R}_m(\mathcal{H} + \mathcal{H}^\prime) = \mathfrak{R}_m(\mathcal{H}) + \mathfrak{R}_m(\mathcal{H^\prime})$.
+\end{enumerate}
+\end{problem}
+
+\begin{solution}
+  \begin{enumerate}[label= (\alph*)]
+    \item We can bound $\mathfrak{R}_m(\alpha \mathcal{H})$ above and below and establish our result using the squeeze theorem.
+          \begin{align*}
+            \mathbb{E}_\sigma \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i (-|\alpha| h(z_i)) \right]                & \leq \mathfrak{R}_m(\alpha \mathcal{H}) & \leq \mathbb{E}_\sigma \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i (|\alpha| h(z_i)) \right] \\
+            \mathbb{E}_{\sigma^\prime} \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i^\prime (|\alpha| h(z_i)) \right] & \leq \mathfrak{R}_m(\alpha \mathcal{H}) & \leq \mathbb{E}_\sigma \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i (|\alpha| h(z_i)) \right]
+          \end{align*}
+          Both sides of the last inequality are equal to each other since $\sigma$ and $\sigma^\prime = -\sigma$ are identically distributed.
+Lastly we use linearity of expectation and the independence of $\alpha$ on $h$ to shift the $|\alpha|$ term outside of the expectation in the following way:
+          \begin{align*}
+            \mathfrak{R}_m(\alpha \mathcal{H}) & = \mathbb{E}_\sigma \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i (|\alpha| h(z_i)) \right] & = |\alpha| \mathbb{E}_\sigma \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i h(z_i) \right] \\
+                                               & = |\alpha| \mathfrak{R}_m(\mathcal{H}) \text{ this concludes the proof}
+          \end{align*}
+          \newpage
+    \item We will rely on the fact that $h$ and $h^\prime$ do not depend on one another so maximising the sum is the same as maximising each separately.
+As a result, we can use the following argument: \begin{align*}
+            \mathfrak{R}_m(\mathcal{H} + \mathcal{H^\prime}) & = \mathbb{E}_\sigma \left[ \sup_{h\in\mathcal{H}, h^\prime \in \mathcal{H}^\prime} \frac{1}{m} \sum_{i=1}^m \sigma_i (h(z_i) + h^\prime (z_i)) \right]                                                                           \\
+                                                             & = \mathbb{E}_\sigma \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i (h(z_i)) \right] + \mathbb{E}_\sigma \left[ \sup_{h^\prime \in \mathcal{H}^\prime} \frac{1}{m} \sum_{i=1}^m \sigma_i (h^\prime(z_i)) \right] \\
+                                                             & = \mathfrak{R}_m(\mathcal{H}) + \mathfrak{R}_m(\mathcal{H}^\prime)
+          \end{align*}
+
+  \end{enumerate}
+\end{solution}
 \section{Convex Learning Problems}

### 2023-06-29 assignment-2/article.tex
Fix whitespace
Finish adding problems for Convex Learning Problems section

@@
 
-\subsection*{Exercise 1.4 (non convexity of 0\-1 loss)}
-Problem 12.1. Hint: Consider samples in a checkerboard pattern.
+\begin{problem}{Exercise 1.4 (non convexity of 0\textendash{}1 loss)}
+Construct an example showing that the 0\textendash{}1 loss function may suffer from local minima; namely, construct a training sample $S \in {(X \times \{\pm1\})}^m$ (say, for $X = \mathbb{R}^2$), for which there exist a vector $w$ and some $\varepsilon > 0$ such that:
+\begin{enumerate}
 
-\subsection*{Exercise 1.5 (Convexity, Lipschitz, and Smoothness of logistic regression loss)}
-Problem 12.2.
+	\item For any $w'$ such that $\|w - w'\| \leq \varepsilon$ we have $L_S (w) \leq L_S (w')$ (where the loss here is the 0\textendash{}1 loss). This means that $w$ is a local minimum of $L_S$.
+	\item There exists some $w^*$ such that $L_S (w^*) < L_S (w)$. This means that $w$ is not a global minimum of $L_S$.
+\end{enumerate}
 
-\subsection*{Exercise 1.6 (Lipschitz continuity of the hinge loss)}
-Problem 12.3.
+Hint: Consider samples in a checkerboard pattern.
+\end{problem}
+
+\begin{solution}
+
+\end{solution}
+
+\begin{problem}{Exercise 1.5 (Convexity, Lipschitz, and Smoothness of logistic regression loss)}
+Consider the learning problem of logistic regression: Let $H = X = \{x \in \mathbb{R}^d : \|x\| \leq B\}$, for some scalar $B > 0$, let $Y = \{\pm1\}$, and let the loss function $\ell$ be defined as $\ell(w, (x, y)) = \log(1 + \exp(-y\langle w, x \rangle))$.
+Show that the resulting learning problem is both convex-Lipschitz-bounded and convex-smooth-bounded.
+Specify the parameters of Lipschitzness and smoothness.
+\end{problem}
+
+\begin{solution}
+
+\end{solution}
+
+\begin{problem}{Exercise 1.6 (Lipschitz continuity of the hinge loss)}
+Consider the problem of learning halfspaces with the hinge loss.
+We limit our domain to the Euclidean ball with radius $R$.
+That is, $X = \{x : \|x\|_2 \leq R\}$.
+The label set is $Y = \{\pm1\}$ and the loss function $\ell$ is defined by $\ell(w, (x, y)) = \max\{0, 1 - y\langle w, x \rangle\}$.
+We already know that the loss function is convex.
+Show that it is $R$-Lipschitz.
+\end{problem}
+
+\begin{solution}
+
+\end{solution}
 

### 2023-06-30 assignment-2/article.tex
Clean up whitespace
First attempt at convex learning

@@
 \begin{solution}
+	Consider three training points in $\mathbb{R}^2$:
+	\[
+		S = \{((0,0), -1), ((0,1), +1), ((1,0), +1)\}
+	\]
+	Let's select a vector $w=(0.5,0.5)$.
+Define the decision boundary as the hyperplane where $w \cdot x = 0.5$.
+This classifier $w$ has a loss of $L_S(w) = \frac{1}{3}$ because it misclassifies the point ((0,0),-1).
 
+	Now, for a small $\varepsilon > 0$, consider the $\varepsilon$-ball around $w$.
+For any vector $w'$ within this ball (i.e., $|w - w'| \leq \varepsilon$), we observe that $L_S(w')$ will not be lower than $L_S(w)$.
+This is because moving the decision boundary within this $\varepsilon$-ball range won't help in correctly classifying the misclassified point.
+Therefore, the vector $w$ is a local minimum of $L_S$.
+
+	However, consider another vector $w^\prime=(1,0)$ which defines the decision boundary as the hyperplane where $w^\prime \cdot x = 0.5$.
+This classifier $w^\prime$ correctly classifies all points in the training set S, and hence, its loss $L_S(w^\prime) = 0$.
+Therefore, we have $L_S(w^\prime) < L_S(w)$, meaning that the vector $w$ is not a global minimum of $L_S$.
+
+	Thus, we have constructed a training sample S for which the 0-1 loss function has a local minimum but not a global minimum. This satisfies all conditions of the problem.
 \end{solution}
@@
 \begin{solution}
+	Given the logistic loss function
+	\[\ell(w, (x, y)) = \log(1 + \exp(-y\langle w, x \rangle))\]
 
+	The gradient of the loss function with respect to $w$ is
+	\[\nabla \ell(w, (x, y)) = -\frac{y \exp(-y\langle w, x \rangle)}{1+\exp(-y\langle w, x \rangle)}x = -\frac{y}{1+\exp(y\langle w, x \rangle)}x\]
+
+	Its norm can be bounded by $B$ because $\|x\| \leq B$. Thus, the loss function is B-Lipschitz.
+
+	To evaluate smoothness, we look at the second derivative of the loss function with respect to $w$. The Hessian is given by
+	\[\nabla^2 \ell(w, (x, y)) = \frac{\exp(y\langle w, x \rangle)}{(1+\exp(y\langle w, x \rangle))^2}xx^T\]
+
+	The spectral norm of the Hessian can be bounded by $B^2$ as $\|x\|^2 \leq B^2$. Thus, the loss function is B²-smooth.
+
+	Therefore, the logistic regression loss function is both B-Lipschitz and B²-smooth.
 \end{solution}
@@
 
+	A function $f: \mathbb{R}^d \rightarrow \mathbb{R}$ is Lipschitz if there exists a constant $L > 0$ such that
+	\[|f(u) - f(v)| \leq L\|u - v\|\]
+	for all $u, v$ in the domain of $f$.
+
+	Here, we are considering the hinge loss function, given by:
+	\[\ell(w, (x, y)) = \max\{0, 1 - y\langle w, x \rangle\}\]
+
+	To show that the hinge loss function is Lipschitz continuous, we have to prove that the difference in loss values for two distinct inputs $w$ and $w'$ is at most $R$ times the distance between $w$ and $w'$.
+
+	Suppose $w, w' \in \mathbb{R}^d$ and $(x, y) \in X \times Y$, then:
+
+	\begin{align*}
+		| \ell(w, (x, y)) - \ell(w', (x, y)) | & = | \max\{0, 1 - y\langle w, x \rangle\} - \max\{0, 1 - y\langle w', x \rangle\} |                                               \\
+		                                       & \leq | y\langle w, x \rangle - y\langle w', x \rangle | \quad \text{(since max function is Lipschitz with Lipschitz constant 1)} \\
+		                                       & = | y\langle w - w', x \rangle |                                                                                                 \\
+		                                       & \leq \| y(x) \|_2 \| w - w' \|_2 \quad \text{(by Cauchy-Schwarz inequality)}                                                     \\
+		                                       & \leq R \| w - w' \|_2 \quad \text{(since } \| x \|_2 \leq R \text{)}
+	\end{align*}
+
+	Therefore, the hinge loss function $\ell$ is $R$-Lipschitz.
 \end{solution}

### 2023-06-30 assignment-2/article.tex
Fix LaTeX formatting

@@
 
-	Thus, we have constructed a training sample S for which the 0-1 loss function has a local minimum but not a global minimum. This satisfies all conditions of the problem.
+	Thus, we have constructed a training sample S for which the 0\textendash{}1 loss function has a local minimum but not a global minimum. This satisfies all conditions of the problem.
 \end{solution}
@@
 	To evaluate smoothness, we look at the second derivative of the loss function with respect to $w$. The Hessian is given by
-	\[\nabla^2 \ell(w, (x, y)) = \frac{\exp(y\langle w, x \rangle)}{(1+\exp(y\langle w, x \rangle))^2}xx^T\]
+	\[\nabla^2 \ell(w, (x, y)) = \frac{\exp(y\langle w, x \rangle)}{(1+{\exp(y\langle w, x \rangle)})^2} x x^T \]
 
@@
 		                                       & \leq \| y(x) \|_2 \| w - w' \|_2 \quad \text{(by Cauchy-Schwarz inequality)}                                                     \\
-		                                       & \leq R \| w - w' \|_2 \quad \text{(since } \| x \|_2 \leq R \text{)}
+		                                       & \leq R \| w - w' \|_2 \quad \text{(since $\| x \|_2 = R$)}
 	\end{align*}

### 2023-07-04 assignment-2/article.tex
Assignment completed

--- 2023-06-29 assignment-2/article.tex
@@
 \begin{solution}
+	Consider three training points in $\mathbb{R}^2$:
+	\[
+		S = \{((0,0), -1), ((0,1), +1), ((1,0), +1)\}
+	\]
+	Let's select a vector $w=(0,0.5)$.
+Define the decision boundary as the hyperplane where $w \cdot x = 0.5$.
+This classifier $w$ has a loss of $L_S(w) = \frac{1}{3}$ because it misclassifies the point ((1,0),+1).
+
+	Let $0 < \varepsilon < \frac{1}{2}$, and consider the $\varepsilon$-ball around $w$.
+For any vector $w'$ within this ball (i.e., $|w - w'| \leq \varepsilon$), we observe that $L_S(w')$ will not be lower than $L_S(w)$.
+This is because in this ball the 0\textendash{}1 loss is constant.
+
+	On the other hand, choosing $w^\prime=(1,1)$ which defines the decision boundary as the hyperplane where $w^\prime \cdot x = 0.5$.
+This classifier $w^\prime$ correctly classifies all points in the training set S, and hence, its loss $L_S(w^\prime) = 0$.
+Therefore, we have $L_S(w^\prime) < L_S(w)$, meaning that the vector $w$ is not a global minimum of $L_S$.
 
@@
 \begin{solution}
+	Given the logistic loss function
+	\[\ell(w, (x, y)) = \log(1 + \exp(-y\langle w, x \rangle))\]
 
+	The magnitude of the gradient of the loss function with respect to $w$ is
+	\begin{align*}
+		||\nabla \ell(w, (x, y))|| & = \left|\left| -\frac{y \exp(-y\langle w, x \rangle)}{1+\exp(-y\langle w, x \rangle)}x \right|\right| \\
+		                           & \leq || -y (1-\sigma(-y\langle w, x\rangle)) \cdot x ||                                               \\
+		                           & \leq ||y|| \cdot |1 - \sigma(-y\langle w,x\rangle)| \cdot ||x||                                       \\
+		                           & \leq B
+	\end{align*}
+
+	Since $||y||=1$, $||x|| \leq B$ and the sigmoid function $\sigma(x)=\frac{1}{1+\exp(-x)}$ is bounded between $[0,1]$. Thus, the loss function is B-Lipschitz.
+
+	To evaluate smoothness, we look at the second derivative of the loss function with respect to $w$. The Hessian is given by
+	\[\nabla^2 \ell(w, (x, y)) = \sigma(-y \langle w,x\rangle) \cdot (1-\sigma(-y \langle w,x\rangle))  x x^T \]
+
+	As before, since the sigmoid is bounded in $[0,1]$ the magnitude of $\nabla^2 \ell(w, (x, y))$ is bounded above by $||xx^T|| \leq B^2$.
+
+	Therefore, the logistic regression loss function is both B-Lipschitz and B²-smooth. Note also that the hessian is positive semi-definite, therefore the logistic regression loss is also convex.
 \end{solution}
@@
 
+	To show that the hinge loss function is Lipschitz continuous, suppose $w, w' \in \mathbb{R}^d$ and $(x, y) \in X \times Y$, then: \begin{align*}
+		| \ell(w, (x, y)) - \ell(w', (x, y)) | & = | \max\{0, 1 - y\langle w, x \rangle\} - \max\{0, 1 - y\langle w', x \rangle\} |                       \\
+		                                       & \leq | y\langle w, x \rangle - y\langle w', x \rangle | \quad \text{(since max is Lipschitz with $L=1$)} \\
+		                                       & = | y\langle w - w', x \rangle |                                                                         \\
+		                                       & \leq \| y(x) \|_2 \| w - w' \|_2 \quad \text{(by Cauchy-Schwarz inequality)}                             \\
+		                                       & \leq R \| w - w' \|_2 \quad \text{(since $\| x \|_2 = R$)}
+	\end{align*}
+
+	Therefore, the hinge loss function $\ell$ is $R$-Lipschitz.
 \end{solution}
